### 2015-06-01 inst/doc/summary.rmd
a few summaries & graphs

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Timings}
-->

```{r opts, echo = FALSE, message = FALSE}
library("knitr")
```

```{r loadpkg,message=FALSE}
library("jsonlite")
library("Timings")
library("plyr") ## handy for ddply, rename
library("ggplot2"); theme_set(theme_bw())
```
```{r}
lf <- list.files(system.file("JSON",package="Timings"),full.names=TRUE)
L1 <- lapply(lf,fromJSON)
sfun <- function(x) {
    ff <- x$models$fits[[1]]
    dd <- data.frame(ds=x$dsname,
               ff[c("deviance","optimizer","function","feval")])
    dd$time <- sapply(ff$time,tail,1)
    ## 'function' and 'deviance' both make trouble for NSE
    dd <- rename(dd,replace=c(function.="fun",deviance="dev"))
    dd <- ddply(dd,"ds",transform,
                ddev=dev-min(dev,na.rm=TRUE),
                reltime=time/min(time,na.rm=TRUE))
    dd
}
## structure of data objects differs ...
has.dsnames <- sapply(lapply(L1,names),function(x) "dsname" %in% x)
sdf <- do.call(rbind,lapply(L1[has.dsnames],sfun))
sdf <- transform(sdf,method=paste(fun,optimizer,sep="."))
sdf <- transform(sdf,method=reorder(method,time),
                     ds=reorder(ds,time))
sdf <- sdf[with(sdf,order(method,ds)),]

## timings
## (we've already ordered methods by average time)
ggplot(sdf,aes(time,method,colour=ds))+
    geom_point()+
    geom_path(aes(group=ds))+
    scale_x_log10()

ggplot(sdf,aes(reltime,method,colour=ds))+
    geom_point()+
    geom_path(aes(group=ds))+
    scale_x_log10()

## taking
ggplot(sdf,aes(ddev,method,colour=ds))+
    geom_point()+
        geom_path(aes(group=ds))+
            geom_vline(xintercept=0.01,linetype=2)+
                scale_x_log10()
ggplot(sdf,aes(ddev,ds,colour=method))+
    geom_point()+
        geom_path(aes(group=method))+
            geom_vline(xintercept=0.01,linetype=2)+
                scale_x_log10()

```

### 2015-06-01 R/retime.R
Use optimx optimizers as well as nloptr.

@@ -23,6 +23,7 @@ retime <- function(fin,fout=fin) {
                               bobyqa = list(maxfun=1e10),
                               Nelder_Mead = list(maxfun=1e10),
                               nloptwrap = list(algorithm=f$algorithm, maxeval=1e10),
+                              nlminbw = list(maxfun=1e10),
                               optimx = list(method=f$method))
             ctrl <- lmerControl(optimizer=f$optimizer,optCtrl=optCtrl)
             tt <- system.time(ff <- lmer(form,dat,REML=FALSE,control=ctrl))
@@ -34,3 +35,5 @@ retime <- function(fin,fout=fin) {
     }
     cat(toJSON(js,digits=I(16),auto_unbox=TRUE,pretty=2),file=fout)
 }
+
+nlminbw <- lme4:::nlminbwrap

### 2015-06-01 R/extractor.R
Add another subject/item data set.  Create an extractor function.

@@ -1,0 +1,17 @@
+revisefr <- function(f) {
+    names(f)[names(f)=="deviance"] <- "dev"
+    f$dev <- round(f$dev,2)
+    names(f)[names(f)=="function"] <- "fun"
+    f$optimizer <- ifelse(is.na(f$algorithm),f$optimizer,f$algorithm)
+    f <- subset(f,select=-c(algorithm,geval))
+    f$time <- sapply(f$time,function(v)v[length(v)])
+    f[order(f$dev,f$time),]
+}
+
+#' Extract information on the model fits from a JSON package as data frames
+#'
+#' @param fin name of the input JSON file
+#' @return a list of data frames.  Each frame is ordered by deviance (rounded to 2 decimal places)
+#'    and by elapsed time withing deviance.
+#' @export
+extractor <- function(fin) lapply(fromJSON(fin)$models$fits,revisefr)

### 2015-06-01 R/extractor.R
Revise the R timing to record only the elapsed time.

@@ -3,7 +3,8 @@ revisefr <- function(f) {
     f$dev <- round(f$dev,2)
     names(f)[names(f)=="function"] <- "fun"
     f$optimizer <- ifelse(is.na(f$algorithm),f$optimizer,f$algorithm)
-    f <- subset(f,select=-c(algorithm,geval))
+    f$optimizer <- ifelse(is.na(f$method), f$optimizer, paste(f$optimizer,f$method,sep=':'))
+    f <- subset(f,select=-c(algorithm,method,geval))
     f$time <- sapply(f$time,function(v)v[length(v)])
     f[order(f$dev,f$time),]
 }

### 2015-06-01 R/retime.R

@@ -27,7 +27,7 @@ retime <- function(fin,fout=fin) {
                               optimx = list(method=f$method))
             ctrl <- lmerControl(optimizer=f$optimizer,optCtrl=optCtrl)
             tt <- system.time(ff <- lmer(form,dat,REML=FALSE,control=ctrl))
-            f$time <- unclass(tt)[1:3]
+            f$time <- unclass(tt)[3]
             f$deviance <- deviance(ff)
             f$feval <- ff@optinfo$feval
             js$models[[i]][["fits"]][[j]] <- f

### 2015-06-02 R/retime.R
Changed maxfun evaluations to eliminate warning messages.

@@ -20,10 +20,10 @@ retime <- function(fin,fout=fin) {
                 next
             opt <- f$optimizer
             optCtrl <- switch(opt,
-                              bobyqa = list(maxfun=1e10),
-                              Nelder_Mead = list(maxfun=1e10),
-                              nloptwrap = list(algorithm=f$algorithm, maxeval=1e10),
-                              nlminbw = list(maxfun=1e10),
+                              bobyqa = list(maxfun=100000),
+                              Nelder_Mead = list(maxfun=100000),
+                              nloptwrap = list(algorithm=f$algorithm, maxeval=100000),
+                              nlminbw = list(maxfun=100000),
                               optimx = list(method=f$method))
             ctrl <- lmerControl(optimizer=f$optimizer,optCtrl=optCtrl)
             tt <- system.time(ff <- lmer(form,dat,REML=FALSE,control=ctrl))

### 2015-07-13 inst/julia/tabulate.jl
Updates to timings and to the tabulate functions.

@@ -33,12 +33,16 @@ function modelsummary(m)
     opt = ASCIIString[]
     times = Float64[]
     devs = Float64[]
+    reltimes = Float64[]
+    basetime = filter(f -> f["optimizer"] == "LN_BOBYQA", m["fits"])[1]["time"]
     for f in m["fits"]
         push!(opt,f["optimizer"])
-        push!(times,f["time"])
+        tt = f["time"]
+        push!(times,tt)
+        push!(reltimes,tt/basetime)
         push!(devs,f["dev"])
     end
-    opt,times,devs,devs .- minimum(devs),fill(m["formula"],length(opt)),
+    opt,times,reltimes,devs,devs .- minimum(devs),fill(m["formula"],length(opt)),
        fill(m["p"],length(opt)),fill(m["q"],length(opt)),fill(m["nopt"],length(opt))
 end
 
@@ -47,6 +51,7 @@ function optdir(dnm)
     models = ASCIIString[]
     opts = ASCIIString[]
     times = Float64[]
+    reltimes = Float64[]
     devs = Float64[]
     excessdev = Float64[]
     n = Int[]
@@ -57,12 +62,13 @@ function optdir(dnm)
         js = JSON.parsefile(joinpath(dnm,nm))
         dsn = js["dsname"]
         for m in js["models"]
-            opt,tt,dev,edevs,forms,pp,qq,nnpp = modelsummary(m)
+            opt,tt,rtt,dev,edevs,forms,pp,qq,nnpp = modelsummary(m)
             append!(models,forms)
             append!(opts,opt)
             append!(devs,dev)
             append!(excessdev,edevs)
             append!(times,tt)
+            append!(reltimes,rtt)
             append!(p,pp)
             append!(q,qq)
             append!(np,nnpp)
@@ -71,7 +77,7 @@ function optdir(dnm)
         end
     end
     ret = DataFrame(opt=pool(opts),dsname=pool(dsnames),n=n,p=p,q=q,np=np,excess=round(excessdev,5),
-                    time=round(times,4),objective=round(devs,5),models=pool(models))
+                    time=round(times,4),reltime=reltimes,objective=round(devs,5),models=pool(models))
     ret[sortperm(ret[:opt]),:]
 end
 

### 2015-07-13 inst/julia/tabulate.jl
Include relative time in summaries.

@@ -77,7 +77,7 @@ function optdir(dnm)
         end
     end
     ret = DataFrame(opt=pool(opts),dsname=pool(dsnames),n=n,p=p,q=q,np=np,excess=round(excessdev,5),
-                    time=round(times,4),reltime=reltimes,objective=round(devs,5),models=pool(models))
+                    time=round(times,4),reltime=round(reltimes,4),objective=round(devs,5),models=pool(models))
     ret[sortperm(ret[:opt]),:]
 end
 

### 2015-07-14 inst/doc/Summary.Rnw
.Rnw file created by 'ipython nbconvert --to latex'

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass{article}

    \usepackage{graphicx} % Used to insert images
    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{color} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2

    \definecolor{orange}{cmyk}{0,0.4,0.8,0.2}
    \definecolor{darkorange}{rgb}{.71,0.21,0.01}
    \definecolor{darkgreen}{rgb}{.12,.54,.11}
    \definecolor{myteal}{rgb}{.26, .44, .56}
    \definecolor{gray}{gray}{0.45}
    \definecolor{lightgray}{gray}{.95}
    \definecolor{mediumgray}{gray}{.8}
    \definecolor{inputbackground}{rgb}{.95, .95, .85}
    \definecolor{outputbackground}{rgb}{.95, .95, .95}
    \definecolor{traceback}{rgb}{1, .95, .95}
    % ansi colors
    \definecolor{red}{rgb}{.6,0,0}
    \definecolor{green}{rgb}{0,.65,0}
    \definecolor{brown}{rgb}{0.6,0.6,0}
    \definecolor{blue}{rgb}{0,.145,.698}
    \definecolor{purple}{rgb}{.698,.145,.698}
    \definecolor{cyan}{rgb}{0,.698,.698}
    \definecolor{lightgray}{gray}{0.5}

    % bright ansi colors
    \definecolor{darkgray}{gray}{0.25}
    \definecolor{lightred}{rgb}{1.0,0.39,0.28}
    \definecolor{lightgreen}{rgb}{0.48,0.99,0.0}
    \definecolor{lightblue}{rgb}{0.53,0.81,0.92}
    \definecolor{lightpurple}{rgb}{0.87,0.63,0.87}
    \definecolor{lightcyan}{rgb}{0.5,1.0,0.83}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{Summaries}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=blue,
      linkcolor=darkorange,
      citecolor=darkgreen,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{Summaries of timing tests}\label{summaries-of-timing-tests}

We compare several optimizers, in both
\href{http://www.R-project.org}{R} and
\href{http://julialang.org}{Julia}, fitting a selection of linear mixed
models. In \textbf{R} the optimizers are called by \texttt{lmer} from
the \href{https://github.com/lme4/lme4}{lme4 package} (version 1.1-8).
In \textbf{Julia} the \texttt{lmm} function from the
\href{https://github.com/dmbates/MixedModels.jl}{MixedModels package}
calls the optimizers.

There are differences in the model formulations in \texttt{lme4} and in
\texttt{MixedModels}. The numerical representation of the model in
\texttt{lme4} and the method of evaluating the objective, described in
\href{http://arxiv.org/abs/1406.5823}{this paper}, is the same for all
models. In \texttt{MixedModels} there are specialized representations
for some model forms, such as models with a single grouping factor for
the random effects. Some of the specialized representations allow for
evaluation of the gradient of the objects, which can enhance convergence
(but, interestingly, sometimes can impede convergence).

\subsection{Methodology}\label{methodology}

To provide consistency we have copied all the data sets used in the
timings to the \texttt{Timings} package itself. We have done all timings
on the same computer. This computer has a relatively recent Intel
processor and we used the
\href{https://software.intel.com/en-us/intel-mkl}{Intel Math Kernel
Library (MKL)} with Julia. We attempted to use
\href{www.revolutionanalytics.com/revolution-r-open}{Revolution R Open
(RRO)} as the R implementation as it can be configured with MKL.
However, we ran into version problems with this so we used the standard
Ubuntu version of R linked against OpenBLAS, which is also
multi-threaded.

Variables were renamed in the pattern: - \textbf{Y} the response -
\textbf{A}, \textbf{B}, $\dots$ categorical covariates - \textbf{G},
\textbf{H}, \textbf{I}, $\dots$ grouping factors for random effects -
\textbf{U}, \textbf{V}, $\dots$ (skipping \textbf{Y}) continuous
covariates

The timing results are saved in \href{http://json.org}{JSON (JavaScript
Object Notation)} files in the directory accessible as

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{system.file}\NormalTok{(}\StringTok{"JSON"}\NormalTok{,}\DataTypeTok{package=}\StringTok{"Timings"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

within \textbf{R}. The directory name will end with
\texttt{./Timings/inst/JSON/} in the package source directory, for
example the result of cloning the
\href{https://github.com/Stat990-033/Timings}{github repository}. There
is one \texttt{.json} file for each data set. Each such file contains
results on timings of one or more models.

The \texttt{Timings} package for \textbf{R} provides a \texttt{retime}
function that takes the name of one of these JSON files and, optionally,
the name of a file with the updated timings. Similarly there are some
source files for Julia retimings.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}1}]:} \PY{n}{include}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../julia/retime.jl}\PY{l+s}{\PYZdq{}}\PY{p}{)}
        \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
        \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
dsname => "Alfalfa"
    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]

        `parse` has no method matching parse(::Int64)
    while loading In[1], in expression starting on line 2

         in retime at /home/bates/git/Timings/inst/julia/retime.jl:17

    \end{Verbatim}

    The timing was repeated so that compilation time is not included in the
results. This repetition is only needed once per session.

A careful examination of these results shows that the main differences
in the Julia timings (the R timings are merely reported, not evaluated)
are that the \texttt{LN\_BOBYQA} and \texttt{LD\_MMA} optimizers are
much faster in the second run. This is because much of the code needs to
be compiled the first time that a derivative-free optimizer and a
derivative-based optimizer are used.

The names of the optimizers used with \texttt{lmm} are those from the
\href{https://github.com/JuliaOpt/NLopt.jl}{NLopt package} for
\textbf{Julia}. Names that begin with \texttt{LD\_} are gradient-based
methods. Names that begin with \texttt{LN\_} are derivative-free
methods. There is one other derivative-free method, \texttt{LN\_PRAXIS},
available in the \texttt{NLopt} package but, for some reason, it can
hang on very simple problems like this. Frequently we omit it.

The optimizers used with \texttt{lmer} include the \texttt{Nelder\_Mead}
optimizer built into the \texttt{lme4} package, the \texttt{bobyqa}
optimizer from the
\href{http://cran.rstudio.com/web/packages/minqa/index.html}{minqa
package}, the derivative-free optimizers from the
\href{http://cran.rstudio.com/web/packages/nloptr/index.html}{nloptr
package} and several optimizers from the
\href{http://cran.rstudio.com/web/packages/optimx/index.html}{optimx
pacakge}.

The \texttt{optimx:bobyqa} optimizer is just a wrapper around
\texttt{bobyqa} (bounded optimization by quadratic approximation) from
the \texttt{minqa} package and should provide results similar to those
from the \texttt{bobyqa} optimizer. For some reason the number of
function evaluations is not reported for the version in \texttt{optimx}.

The optimizers from \texttt{nloptr} (i.e.~those whose names begin with
\texttt{NLOPT\_LN\_}) use the same underlying code as do the similarly
named optimizers in the \texttt{NLopt} package for \textbf{Julia}. The
number of iterations to convergence should be similar for the same
underlying code, although not nessarily exactly the same because the
evaluation of the objective in \textbf{R} and in \textbf{Julia} may
produce slightly different answers. Also the convergence criteria in the
\textbf{Julia} version are more strict than those in the \textbf{R}
version

Also shown are the value of the criterion (negative twice the
log-likelihood, lower is better) achieved, the elapsed time and the
number of function and gradient evaluations. The \texttt{nopt} value is
the number of parameters in the optimization problem. \texttt{mtype} is
the model type in the \textbf{Julia} code. There are special methods for
solving the penalized least squares (PLS) problem, and for evaluating
the objective and its gradient when there is only one grouping factor
for the random effects. The model type is called \texttt{PLSOne}.

The \texttt{Alfalfa} example is a particularly easy one and all of the
optimizerws converge to an objective value close to -10.81023 in less
than 0.6 seconds.

\subsubsection{Tabulating results}\label{tabulating-results}

For the \texttt{Alfalfa} data there is not much of a burden in refitting
the model with all the \textbf{Julia} optimizers just to get the table
shown above. But other examples can take an hour or more to converge and
we don't really need to refit them every time. The \texttt{tabulate.jl}
file contains a function \texttt{optdir} to create a \texttt{DataFrame}
from the results of all the model fits.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}2}]:} \PY{n}{include}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../julia/tabulate.jl}\PY{l+s}{\PYZdq{}}\PY{p}{)}
        \PY{n}{res} \PY{o}{=} \PY{n}{optdir}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
        \PY{n}{res}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{l+m+mi}{30}\PY{p}{,}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{6}\PY{p}{,}\PY{l+m+mi}{7}\PY{p}{,}\PY{l+m+mi}{8}\PY{p}{,}\PY{l+m+mi}{9}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}2}]:} 30x7 DataFrame
        | Row | opt        | dsname          | n      | np | excess  | time    |
        |-----|------------|-----------------|--------|----|---------|---------|
        | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  |
        | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  |
        | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  |
        | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  |
        | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  |
        | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  |
        | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  |
        | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  |
        | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  |
        | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  |
        | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  |
        ⋮
        | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  |
        | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  |
        | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  |
        | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   |
        | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 |
        | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  |
        | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  |
        | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  |
        | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  |
        | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  |
        | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  |
        | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   |

        | Row | reltime |
        |-----|---------|
        | 1   | 1.1342  |
        | 2   | 0.8207  |
        | 3   | 0.9283  |
        | 4   | 0.7897  |
        | 5   | 1.0265  |
        | 6   | 4.4375  |
        | 7   | 0.555   |
        | 8   | 1.3714  |
        | 9   | 0.9925  |
        | 10  | 0.8079  |
        | 11  | 3.9932  |
        ⋮
        | 19  | 0.5604  |
        | 20  | 1.2196  |
        | 21  | 0.4919  |
        | 22  | 0.6573  |
        | 23  | 4.1242  |
        | 24  | 0.6716  |
        | 25  | 4.4016  |
        | 26  | 0.7092  |
        | 27  | 0.9986  |
        | 28  | 4.9414  |
        | 29  | 1.0461  |
        | 30  | 0.9858  |
\end{Verbatim}

    The \texttt{time} column is the time in seconds to converge. The
\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
optimizer in the \texttt{MixedModels} package for \textbf{Julia}.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}3}]:} \PY{n}{res}\PY{p}{[}\PY{n}{res}\PY{p}{[}\PY{p}{:}\PY{n}{opt}\PY{p}{]} \PY{o}{.}\PY{o}{==} \PY{l+s}{\PYZdq{}}\PY{l+s}{NLOPT\PYZus{}LN\PYZus{}BOBYQA}\PY{l+s}{\PYZdq{}}\PY{p}{,}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,}\PY{l+m+mi}{2}\PY{p}{,}\PY{l+m+mi}{3}\PY{p}{,}\PY{l+m+mi}{6}\PY{p}{,}\PY{l+m+mi}{7}\PY{p}{,}\PY{l+m+mi}{8}\PY{p}{,}\PY{l+m+mi}{9}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}3}]:} 49x7 DataFrame
        | Row | opt               | dsname          | n      | np | excess | time   |
        |-----|-------------------|-----------------|--------|----|--------|--------|
        | 1   | "NLOPT\_LN\_BOBYQA" | "Alfalfa"       | 72     | 1  | 0.0    | 0.042  |
        | 2   | "NLOPT\_LN\_BOBYQA" | "Animal"        | 20     | 2  | 0.0    | 0.023  |
        | 3   | "NLOPT\_LN\_BOBYQA" | "Assay"         | 60     | 2  | 1.0e-5 | 0.032  |
        | 4   | "NLOPT\_LN\_BOBYQA" | "AvgDailyGain"  | 32     | 1  | 0.0    | 0.02   |
        | 5   | "NLOPT\_LN\_BOBYQA" | "AvgDailyGain"  | 32     | 1  | 0.0    | 0.02   |
        | 6   | "NLOPT\_LN\_BOBYQA" | "BIB"           | 24     | 1  | 0.0    | 0.02   |
        | 7   | "NLOPT\_LN\_BOBYQA" | "Bond"          | 21     | 1  | 0.0    | 0.02   |
        | 8   | "NLOPT\_LN\_BOBYQA" | "bs10"          | 1104   | 20 | 1.0e-5 | 4.661  |
        | 9   | "NLOPT\_LN\_BOBYQA" | "bs10"          | 1104   | 8  | 0.0    | 1.057  |
        | 10  | "NLOPT\_LN\_BOBYQA" | "cake"          | 270    | 1  | 0.0    | 0.053  |
        | 11  | "NLOPT\_LN\_BOBYQA" | "Chem97"        | 31022  | 2  | 0.0    | 0.632  |
        ⋮
        | 38  | "NLOPT\_LN\_BOBYQA" | "PBIB"          | 60     | 1  | 0.0    | 0.018  |
        | 39  | "NLOPT\_LN\_BOBYQA" | "Penicillin"    | 144    | 2  | 0.0    | 0.023  |
        | 40  | "NLOPT\_LN\_BOBYQA" | "Poems"         | 275996 | 3  | 0.0    | 21.309 |
        | 41  | "NLOPT\_LN\_BOBYQA" | "ScotsSec"      | 3435   | 2  | 0.0    | 0.076  |
        | 42  | "NLOPT\_LN\_BOBYQA" | "Semi2"         | 72     | 3  | 0.0    | 0.03   |
        | 43  | "NLOPT\_LN\_BOBYQA" | "Semiconductor" | 48     | 1  | 0.0    | 0.019  |
        | 44  | "NLOPT\_LN\_BOBYQA" | "SIMS"          | 3691   | 3  | 0.0    | 0.15   |
        | 45  | "NLOPT\_LN\_BOBYQA" | "sleepstudy"    | 180    | 3  | 0.0    | 0.037  |
        | 46  | "NLOPT\_LN\_BOBYQA" | "sleepstudy"    | 180    | 2  | 0.0    | 0.024  |
        | 47  | "NLOPT\_LN\_BOBYQA" | "TeachingII"    | 96     | 1  | 0.0    | 0.021  |
        | 48  | "NLOPT\_LN\_BOBYQA" | "Weights"       | 399    | 3  | 1.0e-5 | 0.039  |
        | 49  | "NLOPT\_LN\_BOBYQA" | "WWheat"        | 60     | 3  | 0.0    | 0.025  |

        | Row | reltime |
        |-----|---------|
        | 1   | 27.5171 |
        | 2   | 13.9236 |
        | 3   | 10.8009 |
        | 4   | 11.7005 |
        | 5   | 13.6969 |
        | 6   | 12.608  |
        | 7   | 23.2699 |
        | 8   | 18.8753 |
        | 9   | 15.6477 |
        | 10  | 22.0827 |
        | 11  | 3.942   |
        ⋮
        | 38  | 13.1358 |
        | 39  | 8.6481  |
        | 40  | 3.7438  |
        | 41  | 5.0422  |
        | 42  | 9.2732  |
        | 43  | 15.9582 |
        | 44  | 1.1034  |
        | 45  | 5.3039  |
        | 46  | 8.6177  |
        | 47  | 15.3763 |
        | 48  | 1.133   |
        | 49  | 2.3154  |
\end{Verbatim}

    \subsection{Proportion converged}\label{proportion-converged}

The most important question regarding the optimizers is whether or not
they have converged to the global optimum. We cannot test this directly.
Instead we use a ``crowd-sourced'' criterion based on the minimum
objective achieved by any of the algorithms. The difference between the
objective achieved by a particular algorithm and this minimum is called
the \texttt{excess}. In the summaries \texttt{excess} is rounded to 5
digits after the decimal so the minimum non-zero \texttt{excess} is
$10^{-5}$.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}4}]:} \PY{n}{res}\PY{p}{[}\PY{n}{res}\PY{p}{[}\PY{p}{:}\PY{n}{opt}\PY{p}{]} \PY{o}{.}\PY{o}{==} \PY{l+s}{\PYZdq{}}\PY{l+s}{LN\PYZus{}BOBYQA}\PY{l+s}{\PYZdq{}}\PY{p}{,}\PY{p}{[}\PY{p}{:}\PY{n}{opt}\PY{p}{,}\PY{p}{:}\PY{n}{dsname}\PY{p}{,}\PY{p}{:}\PY{n}{excess}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}4}]:} 49x3 DataFrame
        | Row | opt         | dsname          | excess |
        |-----|-------------|-----------------|--------|
        | 1   | "LN\_BOBYQA" | "Alfalfa"       | 0.0    |
        | 2   | "LN\_BOBYQA" | "Animal"        | 0.0    |
        | 3   | "LN\_BOBYQA" | "Assay"         | 0.0    |
        | 4   | "LN\_BOBYQA" | "AvgDailyGain"  | 0.0    |
        | 5   | "LN\_BOBYQA" | "AvgDailyGain"  | 0.0    |
        | 6   | "LN\_BOBYQA" | "BIB"           | 0.0    |
        | 7   | "LN\_BOBYQA" | "Bond"          | 0.0    |
        | 8   | "LN\_BOBYQA" | "bs10"          | 1.0e-5 |
        | 9   | "LN\_BOBYQA" | "bs10"          | 0.0    |
        | 10  | "LN\_BOBYQA" | "cake"          | 0.0    |
        | 11  | "LN\_BOBYQA" | "Chem97"        | 0.0    |
        ⋮
        | 38  | "LN\_BOBYQA" | "PBIB"          | 0.0    |
        | 39  | "LN\_BOBYQA" | "Penicillin"    | 0.0    |
        | 40  | "LN\_BOBYQA" | "Poems"         | 0.0    |
        | 41  | "LN\_BOBYQA" | "ScotsSec"      | 0.0    |
        | 42  | "LN\_BOBYQA" | "Semi2"         | 0.0    |
        | 43  | "LN\_BOBYQA" | "Semiconductor" | 0.0    |
        | 44  | "LN\_BOBYQA" | "SIMS"          | 0.0    |
        | 45  | "LN\_BOBYQA" | "sleepstudy"    | 0.0    |
        | 46  | "LN\_BOBYQA" | "sleepstudy"    | 0.0    |
        | 47  | "LN\_BOBYQA" | "TeachingII"    | 0.0    |
        | 48  | "LN\_BOBYQA" | "Weights"       | 0.0    |
        | 49  | "LN\_BOBYQA" | "WWheat"        | 0.0    |
\end{Verbatim}

    If we wish to declare ``converged'' or ``not converged'' according to
the excess objective value we must establish a threshold. An absolute
threshold seems reasonable because the objective, negative twice the
log-likelihood, is on a scale where differences in this objective are
compared to a $\chi^2$ random variable. Thus an excess of $10^{-9}$ or
even $10^{-5}$ is negligible.

For each optimizer we can examine which of the data set/model
combinations resulted in an excess greater than a threshold.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}5}]:} \PY{n}{by}\PY{p}{(}\PY{n}{res}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{)} \PY{k}{do} \PY{n}{df}
            \PY{n}{DataFrame}\PY{p}{(}\PY{n}{attempted}\PY{o}{=}\PY{n}{size}\PY{p}{(}\PY{n}{df}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}\PY{n}{failed}\PY{o}{=}\PY{n}{countnz}\PY{p}{(}\PY{n}{df}\PY{p}{[}\PY{p}{:}\PY{n}{excess}\PY{p}{]} \PY{o}{.}\PY{o}{\PYZgt{}} \PY{l+m+mf}{0.02}\PY{p}{)}\PY{p}{)}
        \PY{k}{end}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}5}]:} 26x3 DataFrame
        | Row | opt                          | attempted | failed |
        |-----|------------------------------|-----------|--------|
        | 1   | "LD\_CCSAQ"                   | 35        | 11     |
        | 2   | "LD\_LBFGS"                   | 35        | 11     |
        | 3   | "LD\_MMA"                     | 36        | 5      |
        | 4   | "LD\_SLSQP"                   | 35        | 4      |
        | 5   | "LD\_TNEWTON"                 | 34        | 8      |
        | 6   | "LD\_TNEWTON\_PRECOND"         | 34        | 8      |
        | 7   | "LD\_TNEWTON\_PRECOND\_RESTART" | 33        | 12     |
        | 8   | "LD\_TNEWTON\_RESTART"         | 34        | 11     |
        | 9   | "LD\_VAR1"                    | 35        | 10     |
        | 10  | "LD\_VAR2"                    | 35        | 10     |
        | 11  | "LN\_BOBYQA"                  | 49        | 0      |
        ⋮
        | 15  | "LN\_SBPLX"                   | 49        | 2      |
        | 16  | "NLOPT\_LN\_BOBYQA"            | 49        | 0      |
        | 17  | "NLOPT\_LN\_COBYLA"            | 48        | 2      |
        | 18  | "NLOPT\_LN\_NELDERMEAD"        | 47        | 6      |
        | 19  | "NLOPT\_LN\_PRAXIS"            | 18        | 4      |
        | 20  | "NLOPT\_LN\_SBPLX"             | 48        | 2      |
        | 21  | "Nelder\_Mead"                | 49        | 8      |
        | 22  | "bobyqa"                     | 49        | 2      |
        | 23  | "optimx:L-BFGS-B"            | 49        | 0      |
        | 24  | "optimx:bobyqa"              | 49        | 2      |
        | 25  | "optimx:nlminb"              | 49        | 0      |
        | 26  | "optimx:spg"                 | 49        | 4      |
\end{Verbatim}

    At this threshold the most reliable algorithm in Julia is
\texttt{LN\_BOBYQA}. In R the most reliable algorithms are
\texttt{NLOPT\_LN\_BOBYQA}, \texttt{optimx:L-BFGS-B} and
\texttt{optimx:nlminb}. It is interesting that nlminb is reliable as I
felt that it wasn't converging well when it was the default optimizer in
\texttt{lmer}.

Interestingly, the derviative-based algorithms in \texttt{NLopt} were
not as reliable as the derivative-free algorithms. The most likely
explanation is that I don't have the gradient coded properly.

The Nelder-Mead simplex algorithm did not perform well, failing on 8 out
of 48 cases. For many of these the value at which convergence was
declared was far from the optimum.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}6}]:} \PY{n}{noncvg} \PY{o}{=} 
        \PY{n}{by}\PY{p}{(}\PY{n}{res}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{)} \PY{k}{do} \PY{n}{df}
            \PY{n}{df}\PY{p}{[}\PY{n}{df}\PY{p}{[}\PY{p}{:}\PY{n}{excess}\PY{p}{]} \PY{o}{.}\PY{o}{\PYZgt{}} \PY{l+m+mf}{0.005}\PY{p}{,}\PY{p}{[}\PY{p}{:}\PY{n}{dsname}\PY{p}{,}\PY{p}{:}\PY{n}{excess}\PY{p}{,}\PY{p}{:}\PY{n}{time}\PY{p}{,}\PY{p}{:}\PY{n}{reltime}\PY{p}{,}\PY{p}{:}\PY{n}{np}\PY{p}{,}\PY{p}{:}\PY{n}{n}\PY{p}{]}\PY{p}{]}
        \PY{k}{end}\PY{p}{;}
        \PY{n}{dfselect}\PY{p}{(}\PY{n}{df}\PY{p}{:}\PY{p}{:}\PY{n}{AbstractDataFrame}\PY{p}{,}\PY{n}{col}\PY{p}{:}\PY{p}{:}\PY{n}{Symbol}\PY{p}{,}\PY{n}{val}\PY{p}{)} \PY{o}{=} \PY{n}{df}\PY{p}{[}\PY{n}{df}\PY{p}{[}\PY{n}{col}\PY{p}{]} \PY{o}{.}\PY{o}{==} \PY{n}{val}\PY{p}{,} \PY{p}{:}\PY{p}{]}
        \PY{n}{dfselect}\PY{p}{(}\PY{n}{noncvg}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{Nelder\PYZus{}Mead}\PY{l+s}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}6}]:} 8x7 DataFrame
        | Row | opt           | dsname        | excess  | time    | reltime | np |
        |-----|---------------|---------------|---------|---------|---------|----|
        | 1   | "Nelder\_Mead" | "bs10"        | 71.3859 | 145.368 | 588.684 | 20 |
        | 2   | "Nelder\_Mead" | "d3"          | 317.59  | 454.519 | 4.2502  | 9  |
        | 3   | "Nelder\_Mead" | "dialectNL"   | 181.632 | 54.541  | 31.1594 | 6  |
        | 4   | "Nelder\_Mead" | "gb12"        | 78.7119 | 38.206  | 189.605 | 20 |
        | 5   | "Nelder\_Mead" | "kb07"        | 398.732 | 2825.46 | 667.015 | 72 |
        | 6   | "Nelder\_Mead" | "kb07"        | 403.478 | 269.436 | 383.087 | 16 |
        | 7   | "Nelder\_Mead" | "Mississippi" | 0.04272 | 0.018   | 20.3989 | 1  |
        | 8   | "Nelder\_Mead" | "mm0"         | 181.632 | 76.87   | 70.0716 | 6  |

        | Row | n      |
        |-----|--------|
        | 1   | 1104   |
        | 2   | 130418 |
        | 3   | 225866 |
        | 4   | 512    |
        | 5   | 1790   |
        | 6   | 1790   |
        | 7   | 37     |
        | 8   | 69588  |
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}7}]:} \PY{n}{dfselect}\PY{p}{(}\PY{n}{noncvg}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{NLOPT\PYZus{}LN\PYZus{}NELDERMEAD}\PY{l+s}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}7}]:} 6x7 DataFrame
        | Row | opt                   | dsname  | excess  | time    | reltime | np |
        |-----|-----------------------|---------|---------|---------|---------|----|
        | 1   | "NLOPT\_LN\_NELDERMEAD" | "Assay" | 0.05942 | 0.042   | 14.1762 | 2  |
        | 2   | "NLOPT\_LN\_NELDERMEAD" | "bs10"  | 0.97096 | 88.286  | 357.524 | 20 |
        | 3   | "NLOPT\_LN\_NELDERMEAD" | "d3"    | 100.191 | 627.931 | 5.8718  | 9  |
        | 4   | "NLOPT\_LN\_NELDERMEAD" | "gb12"  | 0.83883 | 52.411  | 260.1   | 20 |
        | 5   | "NLOPT\_LN\_NELDERMEAD" | "kb07"  | 88.7686 | 2839.07 | 670.229 | 72 |
        | 6   | "NLOPT\_LN\_NELDERMEAD" | "kb07"  | 3.49868 | 198.845 | 282.72  | 16 |

        | Row | n      |
        |-----|--------|
        | 1   | 60     |
        | 2   | 1104   |
        | 3   | 130418 |
        | 4   | 512    |
        | 5   | 1790   |
        | 6   | 1790   |
\end{Verbatim}

    The \texttt{Nelder\_Mead} algorithm, either in the native form in
\texttt{lmer} or in the \texttt{NLopt} implementation performed poorly
on those cases with many parameters to optimize. It was both unreliable
and slow, taking over 45 minutes to reach a spurious optimum on the
``maximal'' model (in the sense of Barr et al., 2012) for the
\texttt{kb07} data from Kronmueller and Barr (2007). This is not
terribly surprising given that the model is horribly overparameterized,
but still it shows that this algorithm is not a good choice in these
cases.

We note in passing that all the models involving fitting 20 or more
parameters are ``maximal'' models in the sense of Barr et al., 2012.
Such models can present difficult optimization problems because they are
severely overparameterized and inevitably converge on the boundary of
the allowable parameter space. Whether or not it is sensible to compare
results on such extreme cases is not clear.

The \texttt{SBPLX} (subplex) algorithm, which is an enhancement of
\texttt{Nelder\_Mead}, does better in these cases but is still rather
slow.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}8}]:} \PY{n}{dfselect}\PY{p}{(}\PY{n}{noncvg}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{NLOPT\PYZus{}LN\PYZus{}SBPLX}\PY{l+s}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}8}]:} 2x7 DataFrame
        | Row | opt              | dsname | excess  | time    | reltime | np | n    |
        |-----|------------------|--------|---------|---------|---------|----|------|
        | 1   | "NLOPT\_LN\_SBPLX" | "gb12" | 0.82219 | 3.813   | 18.9228 | 20 | 512  |
        | 2   | "NLOPT\_LN\_SBPLX" | "kb07" | 4.96546 | 564.688 | 133.308 | 72 | 1790 |
\end{Verbatim}

    By comparison, the \texttt{LN\_BOBYQA} algorithm converges quite rapidly
on the \texttt{kb07} models.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}9}]:} \PY{n}{bobyqa} \PY{o}{=} \PY{n}{res}\PY{p}{[}\PY{n+nb}{convert}\PY{p}{(}\PY{n}{Array}\PY{p}{,}\PY{n}{res}\PY{p}{[}\PY{p}{:}\PY{n}{opt}\PY{p}{]} \PY{o}{.}\PY{o}{==} \PY{l+s}{\PYZdq{}}\PY{l+s}{LN\PYZus{}BOBYQA}\PY{l+s}{\PYZdq{}}\PY{p}{)} \PY{o}{\PYZam{}} 
            \PY{n+nb}{convert}\PY{p}{(}\PY{n}{Array}\PY{p}{,}\PY{n}{res}\PY{p}{[}\PY{p}{:}\PY{n}{dsname}\PY{p}{]} \PY{o}{.}\PY{o}{==} \PY{l+s}{\PYZdq{}}\PY{l+s}{kb07}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{,}
            \PY{p}{[}\PY{p}{:}\PY{n}{dsname}\PY{p}{,}\PY{p}{:}\PY{n}{excess}\PY{p}{,}\PY{p}{:}\PY{n}{objective}\PY{p}{,}\PY{p}{:}\PY{n}{time}\PY{p}{,}\PY{p}{:}\PY{n}{np}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}9}]:} 2x5 DataFrame
        | Row | dsname | excess  | objective | time   | np |
        |-----|--------|---------|-----------|--------|----|
        | 1   | "kb07" | 0.01695 | 28586.3   | 4.236  | 72 |
        | 2   | "kb07" | 0.0     | 28670.9   | 0.7033 | 16 |
\end{Verbatim}

    \subsection{Relative speed}\label{relative-speed}

We plot the time to convergence, relative to \texttt{LN\_BOBYQA} and on
a logarithmic scale, for each algorithm.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}10}]:} \PY{k}{using} \PY{n}{Gadfly}
         \PY{n}{set\PYZus{}default\PYZus{}plot\PYZus{}size}\PY{p}{(}\PY{l+m+mi}{16}\PY{n}{cm}\PY{p}{,}\PY{l+m+mi}{12}\PY{n}{cm}\PY{p}{)}
         \PY{n}{res}\PY{p}{[}\PY{p}{:}\PY{n}{cvg}\PY{p}{]} \PY{o}{=} \PY{n}{compact}\PY{p}{(}\PY{n}{pool}\PY{p}{(}\PY{p}{[}\PY{n}{e} \PY{o}{\PYZgt{}} \PY{l+m+mf}{0.02} \PY{o}{?} \PY{l+s}{\PYZdq{}}\PY{l+s}{N}\PY{l+s}{\PYZdq{}} \PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Y}\PY{l+s}{\PYZdq{}} \PY{k}{for} \PY{n}{e} \PY{k}{in} \PY{n}{res}\PY{p}{[}\PY{p}{:}\PY{n}{excess}\PY{p}{]}\PY{p}{]}\PY{p}{)}\PY{p}{)}
         \PY{n}{plot}\PY{p}{(}\PY{n}{res}\PY{p}{,}\PY{n}{x}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{reltime}\PY{l+s}{\PYZdq{}}\PY{p}{,}\PY{n}{y}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{opt}\PY{l+s}{\PYZdq{}}\PY{p}{,}\PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{cvg}\PY{l+s}{\PYZdq{}}\PY{p}{,}
             \PY{n}{Geom}\PY{o}{.}\PY{n}{point}\PY{p}{,}\PY{n}{Scale}\PY{o}{.}\PY{n}{y\PYZus{}discrete}\PY{p}{,}\PY{n}{Scale}\PY{o}{.}\PY{n}{x\PYZus{}log2}\PY{p}{,}
             \PY{n}{Guide}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{n}{nothing}\PY{p}{)}\PY{p}{,}
             \PY{n}{Guide}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Time to convergence relative to LN\PYZus{}BOBYQA}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\texttt{\color{outcolor}Out[{\color{outcolor}10}]:}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Summaries_files/Summaries_18_0.pdf}
    \end{center}
    { \hspace*{\fill} \\}

    Many of the cases where \texttt{LN\_BOBYQA} is slower than other
algorithms are simple problems that converge in less than 1/5 of a
second for most algorithms.

We will declare a data set to be non-simple if at least one of the
models fit to the data took more than 0.2 seconds to convergence with
\texttt{LN\_BOBYQA}.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}12}]:} \PY{n}{ln\PYZus{}bobyqa} \PY{o}{=} \PY{n}{dfselect}\PY{p}{(}\PY{n}{res}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{LN\PYZus{}BOBYQA}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
         \PY{n}{nonsimple} \PY{o}{=} \PY{n}{ln\PYZus{}bobyqa}\PY{p}{[}\PY{n}{ln\PYZus{}bobyqa}\PY{p}{[}\PY{p}{:}\PY{n}{time}\PY{p}{]} \PY{o}{.}\PY{o}{\PYZgt{}} \PY{l+m+mf}{0.2}\PY{p}{,}\PY{p}{[}\PY{p}{:}\PY{n}{dsname}\PY{p}{,}\PY{p}{:}\PY{n}{time}\PY{p}{,}\PY{p}{:}\PY{n}{n}\PY{p}{,}\PY{p}{:}\PY{n}{np}\PY{p}{,}\PY{p}{:}\PY{n}{models}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}12}]:} 10x5 DataFrame
         | Row | dsname      | time   | n      | np |
         |-----|-------------|--------|--------|----|
         | 1   | "bs10"      | 0.2469 | 1104   | 20 |
         | 2   | "d3"        | 106.94 | 130418 | 9  |
         | 3   | "dialectNL" | 1.7504 | 225866 | 6  |
         | 4   | "gb12"      | 0.2015 | 512    | 20 |
         | 5   | "InstEval"  | 2.3539 | 73421  | 2  |
         | 6   | "InstEval"  | 4.3593 | 73421  | 3  |
         | 7   | "kb07"      | 4.236  | 1790   | 72 |
         | 8   | "kb07"      | 0.7033 | 1790   | 16 |
         | 9   | "mm0"       | 1.097  | 69588  | 6  |
         | 10  | "Poems"     | 5.6918 | 275996 | 3  |

         | Row | models                                                                                                                                                |
         |-----|-------------------------------------------------------------------------------------------------------------------------------------------------------|
         | 1   | "29"                                                                                                                                                  |
         | 2   | "Y \textasciitilde{} U + (U | G)  + (U | H) + (U | I)"                                                                                                                |
         | 3   | "Y \textasciitilde{} U + V +W +X + Z+A + T +(1 | G) +(0 + V + W+ X | G) +(1 + Z + A | H) +(1 | I)"                                                                    |
         | 4   | "Y \textasciitilde{} 1+S+T+U+V+W+X+Z + (1+S+U+W|G) + (1+S+T+V|H)"                                                                                                     |
         | 5   | "Y \textasciitilde{} 1 + I*A + (1|G) + (1|H)"                                                                                                                         |
         | 6   | "Y \textasciitilde{} 1 + A + (1|G) + (1|H) + (1|I)"                                                                                                                   |
         | 7   | "Y \textasciitilde{} 1+S+T+U+V+W+X+Z + (1+S+T+U+V+W+X+Z|G) + (1+S+T+U+V+W+X+Z|H)"                                                                                     |
         | 8   | "Y \textasciitilde{} 1+S+T+U+V+W+X+Z + (1|G)+(0+S|G)+(0+T|G)+(0+U|G)+(0+V|G)+(0+W|G)+(0+X|G)+(0+Z|G) + (1|H)+(0+S|H)+(0+T|H)+(0+U|H)+(0+V|H)+(0+W|H)+(0+X|H)+(0+Z|H)" |
         | 9   | "Y\textasciitilde{}A*U+(1+U|G)+(1+U|H)"                                                                                                                               |
         | 10  | "Y \textasciitilde{} 1 + U + V + (1|G) + (1|H) + (1|I)"                                                                                                               |
\end{Verbatim}

    Notice that these are cases with a large number of observations
(\texttt{n}) or a large number of parameters in the optimization problem
(\texttt{np}) or both.

By comparison, the cases where other algorithms are faster than
\texttt{LN\_BOBYQA} are, for the most part, models and data sets with
few observations and few parameters to optimize. In these circumstances
almost all the optimizers are fast.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}13}]:} \PY{n}{by}\PY{p}{(}\PY{n}{res}\PY{p}{,}\PY{p}{:}\PY{n}{opt}\PY{p}{)} \PY{k}{do} \PY{n}{df}
             \PY{n}{cvgfast} \PY{o}{=} \PY{p}{(}\PY{n}{df}\PY{p}{[}\PY{p}{:}\PY{n}{reltime}\PY{p}{]} \PY{o}{.}\PY{o}{\PYZlt{}} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZam{}} \PY{n+nb}{convert}\PY{p}{(}\PY{n}{Array}\PY{p}{,}\PY{n}{df}\PY{p}{[}\PY{p}{:}\PY{n}{cvg}\PY{p}{]} \PY{o}{.}\PY{o}{==} \PY{l+s}{\PYZdq{}}\PY{l+s}{Y}\PY{l+s}{\PYZdq{}}\PY{p}{)}
             \PY{k}{if} \PY{n}{any}\PY{p}{(}\PY{n}{cvgfast}\PY{p}{)}
                 \PY{n}{print}\PY{p}{(}\PY{n}{df}\PY{p}{[}\PY{n}{cvgfast}\PY{p}{,}\PY{p}{[}\PY{p}{:}\PY{n}{opt}\PY{p}{,}\PY{p}{:}\PY{n}{dsname}\PY{p}{,}\PY{p}{:}\PY{n}{time}\PY{p}{,}\PY{p}{:}\PY{n}{np}\PY{p}{,}\PY{p}{:}\PY{n}{n}\PY{p}{]}\PY{p}{]}\PY{p}{)}
                 \PY{n}{println}\PY{p}{(}\PY{p}{)}
             \PY{k}{end}
         \PY{k}{end}\PY{p}{;}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
8x5 DataFrame
| Row | opt        | dsname         | time   | np | n  |
|-----|------------|----------------|--------|----|----|
| 1   | "LD\_CCSAQ" | "AvgDailyGain" | 0.0014 | 1  | 32 |
| 2   | "LD\_CCSAQ" | "AvgDailyGain" | 0.0014 | 1  | 32 |
| 3   | "LD\_CCSAQ" | "BIB"          | 0.0013 | 1  | 24 |
| 4   | "LD\_CCSAQ" | "Cultivation"  | 0.0009 | 1  | 24 |
| 5   | "LD\_CCSAQ" | "Dyestuff2"    | 0.0005 | 1  | 30 |
| 6   | "LD\_CCSAQ" | "Gasoline"     | 0.0011 | 1  | 32 |
| 7   | "LD\_CCSAQ" | "PBIB"         | 0.0014 | 1  | 60 |
| 8   | "LD\_CCSAQ" | "TeachingII"   | 0.0013 | 1  | 96 |
2x5 DataFrame
| Row | opt        | dsname | time   | np | n   |
|-----|------------|--------|--------|----|-----|
| 1   | "LD\_LBFGS" | "gb12" | 0.1871 | 20 | 512 |
| 2   | "LD\_LBFGS" | "HR"   | 0.004  | 3  | 120 |
12x5 DataFrame
| Row | opt      | dsname         | time   | np | n   |
|-----|----------|----------------|--------|----|-----|
| 1   | "LD\_MMA" | "AvgDailyGain" | 0.0015 | 1  | 32  |
| 2   | "LD\_MMA" | "BIB"          | 0.0012 | 1  | 24  |
| 3   | "LD\_MMA" | "Bond"         | 0.0008 | 1  | 21  |
| 4   | "LD\_MMA" | "Dyestuff2"    | 0.0005 | 1  | 30  |
| 5   | "LD\_MMA" | "Dyestuff"     | 0.0007 | 1  | 30  |
| 6   | "LD\_MMA" | "ergoStool"    | 0.0009 | 1  | 36  |
| 7   | "LD\_MMA" | "Gasoline"     | 0.0012 | 1  | 32  |
| 8   | "LD\_MMA" | "Oxboys"       | 0.0092 | 3  | 234 |
| 9   | "LD\_MMA" | "PBIB"         | 0.0013 | 1  | 60  |
| 10  | "LD\_MMA" | "TeachingII"   | 0.0012 | 1  | 96  |
| 11  | "LD\_MMA" | "Weights"      | 0.027  | 3  | 399 |
| 12  | "LD\_MMA" | "WWheat"       | 0.0084 | 3  | 60  |
20x5 DataFrame
| Row | opt        | dsname          | time   | np | n    |
|-----|------------|-----------------|--------|----|------|
| 1   | "LD\_SLSQP" | "Alfalfa"       | 0.0014 | 1  | 72   |
| 2   | "LD\_SLSQP" | "AvgDailyGain"  | 0.0016 | 1  | 32   |
| 3   | "LD\_SLSQP" | "BIB"           | 0.0011 | 1  | 24   |
| 4   | "LD\_SLSQP" | "Bond"          | 0.0007 | 1  | 21   |
| 5   | "LD\_SLSQP" | "Cultivation"   | 0.0008 | 1  | 24   |
| 6   | "LD\_SLSQP" | "Demand"        | 0.0054 | 2  | 77   |
| 7   | "LD\_SLSQP" | "Dyestuff2"     | 0.0007 | 1  | 30   |
| 8   | "LD\_SLSQP" | "Dyestuff"      | 0.0007 | 1  | 30   |
| 9   | "LD\_SLSQP" | "ergoStool"     | 0.001  | 1  | 36   |
| 10  | "LD\_SLSQP" | "Gasoline"      | 0.0011 | 1  | 32   |
| 11  | "LD\_SLSQP" | "HR"            | 0.0056 | 3  | 120  |
| 12  | "LD\_SLSQP" | "IncBlk"        | 0.0013 | 1  | 24   |
| 13  | "LD\_SLSQP" | "Oxboys"        | 0.0072 | 3  | 234  |
| 14  | "LD\_SLSQP" | "PBIB"          | 0.0009 | 1  | 60   |
| 15  | "LD\_SLSQP" | "Semiconductor" | 0.001  | 1  | 48   |
| 16  | "LD\_SLSQP" | "SIMS"          | 0.0742 | 3  | 3691 |
| 17  | "LD\_SLSQP" | "sleepstudy"    | 0.0052 | 3  | 180  |
| 18  | "LD\_SLSQP" | "TeachingII"    | 0.0011 | 1  | 96   |
| 19  | "LD\_SLSQP" | "Weights"       | 0.0234 | 3  | 399  |
| 20  | "LD\_SLSQP" | "WWheat"        | 0.0079 | 3  | 60   |
3x5 DataFrame
| Row | opt          | dsname    | time   | np | n   |
|-----|--------------|-----------|--------|----|-----|
| 1   | "LD\_TNEWTON" | "HR"      | 0.0059 | 3  | 120 |
| 2   | "LD\_TNEWTON" | "Oxboys"  | 0.0087 | 3  | 234 |
| 3   | "LD\_TNEWTON" | "Weights" | 0.0213 | 3  | 399 |
3x5 DataFrame
| Row | opt                  | dsname    | time   | np | n   |
|-----|----------------------|-----------|--------|----|-----|
| 1   | "LD\_TNEWTON\_PRECOND" | "HR"      | 0.0048 | 3  | 120 |
| 2   | "LD\_TNEWTON\_PRECOND" | "Oxboys"  | 0.0092 | 3  | 234 |
| 3   | "LD\_TNEWTON\_PRECOND" | "Weights" | 0.0152 | 3  | 399 |
1x5 DataFrame
| Row | opt                          | dsname | time  | np | n   |
|-----|------------------------------|--------|-------|----|-----|
| 1   | "LD\_TNEWTON\_PRECOND\_RESTART" | "HR"   | 0.005 | 3  | 120 |
1x5 DataFrame
| Row | opt                  | dsname | time   | np | n   |
|-----|----------------------|--------|--------|----|-----|
| 1   | "LD\_TNEWTON\_RESTART" | "HR"   | 0.0052 | 3  | 120 |
1x5 DataFrame
| Row | opt       | dsname | time   | np | n   |
|-----|-----------|--------|--------|----|-----|
| 1   | "LD\_VAR1" | "HR"   | 0.0044 | 3  | 120 |
1x5 DataFrame
| Row | opt       | dsname | time   | np | n   |
|-----|-----------|--------|--------|----|-----|
| 1   | "LD\_VAR2" | "HR"   | 0.0044 | 3  | 120 |
2x5 DataFrame
| Row | opt         | dsname  | time   | np | n    |
|-----|-------------|---------|--------|----|------|
| 1   | "LN\_COBYLA" | "BIB"   | 0.0014 | 1  | 24   |
| 2   | "LN\_COBYLA" | "Hsb82" | 0.018  | 1  | 7185 |
3x5 DataFrame
| Row | opt             | dsname      | time   | np | n   |
|-----|-----------------|-------------|--------|----|-----|
| 1   | "LN\_NELDERMEAD" | "BIB"       | 0.0015 | 1  | 24  |
| 2   | "LN\_NELDERMEAD" | "Dyestuff2" | 0.0005 | 1  | 30  |
| 3   | "LN\_NELDERMEAD" | "Oxboys"    | 0.0192 | 3  | 234 |
1x5 DataFrame
| Row | opt         | dsname  | time   | np | n      |
|-----|-------------|---------|--------|----|--------|
| 1   | "LN\_PRAXIS" | "Poems" | 5.3089 | 3  | 275996 |
4x5 DataFrame
| Row | opt        | dsname      | time    | np | n      |
|-----|------------|-------------|---------|----|--------|
| 1   | "LN\_SBPLX" | "d3"        | 75.8816 | 9  | 130418 |
| 2   | "LN\_SBPLX" | "Demand"    | 0.0049  | 2  | 77     |
| 3   | "LN\_SBPLX" | "dialectNL" | 1.4855  | 6  | 225866 |
| 4   | "LN\_SBPLX" | "Dyestuff2" | 0.0006  | 1  | 30     |
1x5 DataFrame
| Row | opt               | dsname | time  | np | n    |
|-----|-------------------|--------|-------|----|------|
| 1   | "NLOPT\_LN\_COBYLA" | "SIMS" | 0.092 | 3  | 3691 |
    \end{Verbatim}

    % Add a bibliography block to the postdoc

    \end{document}

### 2015-07-30 inst/doc/Paper.Rnw
Added necessary files for new rnw file

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Achim Zeileis\\Universit\"at Innsbruck \And 
        Second Author\\Plus Affiliation}
\title{A Capitalized Title: Something about Package \pkg{Timings}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Achim Zeileis, Second Author} %% comma-separated
\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
\Shorttitle{\pkg{foo}: A Capitalized Title} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  The abstract of the article.
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Achim Zeileis\\
  Department of Statistics and Mathematics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@uibk.ac.at}\\
  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\SweaveOpts{concordance=FALSE}
%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}

### 2015-07-30 inst/doc/Paper.tex
Basic edits

--- 2015-07-30 inst/doc/Paper.Rnw
@@
 %% almost as usual
-\author{Achim Zeileis\\Universit\"at Innsbruck \And 
-        Second Author\\Plus Affiliation}
-\title{A Capitalized Title: Something about Package \pkg{Timings}}
+\author{Douglas Bates\\University\ of Wisconsin-Madison \And 
+        Colin Longhurst \\University of Wisconsin-Madison}
+\title{Comparing Optimization Algorithms in the Fitting of Linear Mixed Models: Evaluating Speed and Accuracy using \pkg{lme4} in R and \pkg{lmm} in Julia}
 
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Achim Zeileis, Second Author} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Package foo} %% without formatting
-\Shorttitle{\pkg{foo}: A Capitalized Title} %% a short title (if necessary)
+\Plainauthor{Douglas Bates, Colin Longhurst} %% comma-separated
+\Plaintitle{Comparing Optimization Algorithms in the Fitting of Linear Mixed Models: Evaluating Speed and Accuracy using lme4 in R and lmm in Julia} %% without formatting
+\Shorttitle{\pkg{Timings}: A Package for the Evaluation of Optimization Algorithms} %% a short title (if necessary)
 
@@
 \Abstract{
-  The abstract of the article.
+  Donec pulvinar ligula vel sem finibus, vitae pulvinar metus tincidunt.
+Nam at arcu sem.
+Aenean in leo viverra, consectetur ex nec, malesuada lacus.
+Integer venenatis porta ullamcorper.
+Cras viverra lectus commodo porttitor dictum.
+Sed quis nulla rhoncus metus pretium vestibulum.
+Quisque ut scelerisque tellus, non egestas risus.
+Quisque finibus ligula id sem ultrices convallis.
+Vivamus maximus placerat nunc, ac posuere eros viverra quis.
+Donec magna urna, gravida in orci sed, sagittis congue magna.
+Duis nec nibh vel augue interdum iaculis a sit amet ipsum.
+Nunc lectus quam, varius sed sodales in, iaculis et diam.
 }
-\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
-\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
+\Keywords{optimizers, mixed models, linear mixed models, lme4, lmm, \proglang{R}, \proglang{Julia}}
+\Plainkeywords{optimizers, mixed models, linear mixed models, lme4, lmm} %% without formatting
 %% at least one keyword must be supplied
@@
 \begin{document}
-\SweaveOpts{concordance=FALSE}
+
 %% include your article here, just as usual

### 2015-07-31 inst/doc/Paper.Rnw
more edits

@@
 \Address{
-  Achim Zeileis\\
-  Department of Statistics and Mathematics\\
-  Faculty of Economics and Statistics\\
-  Universit\"at Innsbruck\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{Achim.Zeileis@uibk.ac.at}\\
-  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
+  Douglas Bates\\
+  Emeritus Professor\\
+  Department of Statistics\\
+  University of Wisconsin-Madison\\
+  1300 University Ave
+  Madison, WI 53706-1685
+  U.S.A.\\
+  E-mail: \email{dmbates@stat.wisc.edu}\\
+  Github URL: \url{https://github.com/dmbates}
 }
@@
 
-\section[About Java]{About \proglang{Java}}
+\section[Introduction]{Introduction}
 %% Note: If there is markup in \(sub)section, then it has to be escape as above.
 
+Mauris ac lectus sagittis, mollis diam porttitor, efficitur purus.
+Duis ut eros vitae dolor vulputate sagittis.
+Aenean viverra, dui in interdum consectetur, lectus elit placerat urna, dictum ultricies ex justo eget est.
+In at leo quam.
+Suspendisse eros nulla, gravida nec suscipit vel, egestas eget neque.
+Etiam pharetra, nisl nec viverra ultricies, mauris lectus facilisis velit, vel aliquam ipsum lorem ut sem.
+Nulla ornare, justo at tempus blandit, nulla urna porta ligula, id dictum mi est ac massa.
+Quisque at lacus neque.
+Pellentesque est nulla, dignissim eget lobortis ut, vehicula at erat.
+Aenean ornare lacus mattis, elementum elit vel, tempor risus.
+In in purus tempor lacus imperdiet rhoncus nec nec tortor.
+Duis sagittis nisl ante, id egestas neque tristique fermentum.
+Fusce aliquet, odio non auctor aliquet, purus orci venenatis purus, sit amet pulvinar nisi est at dolor.
+Pellentesque lobortis dui eros, et ultricies tellus ultrices at.
+Ut sit amet interdum justo.
+Integer placerat vehicula interdum.
+
 \end{document}

### 2015-07-31 inst/doc/Paper.Rnw
More edits, added sections, dummy text etc.

@@
 \Abstract{
-  Donec pulvinar ligula vel sem finibus, vitae pulvinar metus tincidunt.
-Nam at arcu sem.
-Aenean in leo viverra, consectetur ex nec, malesuada lacus.
-Integer venenatis porta ullamcorper.
-Cras viverra lectus commodo porttitor dictum.
-Sed quis nulla rhoncus metus pretium vestibulum.
-Quisque ut scelerisque tellus, non egestas risus.
-Quisque finibus ligula id sem ultrices convallis.
-Vivamus maximus placerat nunc, ac posuere eros viverra quis.
-Donec magna urna, gravida in orci sed, sagittis congue magna.
-Duis nec nibh vel augue interdum iaculis a sit amet ipsum.
-Nunc lectus quam, varius sed sodales in, iaculis et diam.
+The \pkg{Timings} package allows for the comparison of several optimizers, in both \proglang{R} and \proglang{Julia}, used in the fitting of various linear mixed models.
+In \proglang{R} the optimizers are called by lmer from the \pkg{lme4} package (version 1.1-8).
+In \proglang{Julia} the optimizers are called by lmm from \pkg{MixedModels} package.
+From the \pkg{Timings} package, conclusions regarding an optimizers relative speed, accuracy and general effectiveness of different optimizers paired with different types of models (ranging from simple to complex) can easily be drawn and interpretted.
+\\
+  There are differences in the model formulations in \pkg{lme4} and in \pkg{MixedModels}.
+The numerical representation of the model in \pkg{lme4} and the method of evaluating the optimizers, described in this paper, is the same for all models.
+In \pkg{MixedModels} there are specialized representations for some model forms, such as models with a single grouping factor for the random effects.
+Some of the specialized representations allow for evaluation of the gradient of the objects, which can enhance convergence (but, interestingly, sometimes can impede convergence).
 }
@@
 
+\section[Methods]{Methods}
+
+Mauris ac lectus sagittis, mollis diam porttitor, efficitur purus.
+Duis ut eros vitae dolor vulputate sagittis.
+Aenean viverra, dui in interdum consectetur, lectus elit placerat urna, dictum ultricies ex justo eget est.
+In at leo quam.
+Suspendisse eros nulla, gravida nec suscipit vel, egestas eget neque.
+Etiam pharetra, nisl nec viverra ultricies, mauris lectus facilisis velit, vel aliquam ipsum lorem ut sem.
+Nulla ornare, justo at tempus blandit, nulla urna porta ligula, id dictum mi est ac massa.
+Quisque at lacus neque.
+Pellentesque est nulla, dignissim eget lobortis ut, vehicula at erat.
+Aenean ornare lacus mattis, elementum elit vel, tempor risus.
+In in purus tempor lacus imperdiet rhoncus nec nec tortor.
+Duis sagittis nisl ante, id egestas neque tristique fermentum.
+Fusce aliquet, odio non auctor aliquet, purus orci venenatis purus, sit amet pulvinar nisi est at dolor.
+Pellentesque lobortis dui eros, et ultricies tellus ultrices at.
+Ut sit amet interdum justo.
+Integer placerat vehicula interdum.
+
+\section[Results]{Results}
+
+Nullam eget gravida justo.
+Proin lectus mauris, dictum vitae ultricies ut, accumsan sed nisi.
+Pellentesque quis nunc vestibulum, aliquam arcu eu, dapibus risus.
+Pellentesque lobortis nulla at posuere molestie.
+Etiam risus nisi, pellentesque a tortor at, porta mollis est.
+Curabitur nec ante eget felis pellentesque sagittis id in nulla.
+Vivamus sit amet accumsan lorem.
+Nam egestas ultrices nisi, volutpat rutrum tortor tincidunt ac.
+Ut ex nunc, ornare eget nibh quis, mollis aliquam dui.
+Sed tempor, sapien sed vestibulum malesuada, dolor est ullamcorper leo, at mollis erat nulla eget urna.
+Aenean risus leo, pellentesque sit amet pulvinar a, ultrices ut lacus.
+Sed id nunc tempus, commodo augue a, hendrerit nunc.
+Mauris et purus lorem.
+Praesent ornare, ipsum id bibendum placerat, arcu enim iaculis tellus, tincidunt rhoncus augue dui sed ex.
+Proin nunc nisl, lacinia quis tristique eget, tempor eu velit.
+
+\subsection[Speed]{Speed}
+ Etiam placerat commodo dapibus.
+Vivamus a porta dui.
+Phasellus accumsan sollicitudin cursus.
+In nec lorem sit amet purus vulputate suscipit quis ac lectus.
+Aliquam sollicitudin malesuada blandit.
+Aliquam sed nisi accumsan, pulvinar diam eget, facilisis urna.
+Fusce a dignissim ipsum.
+Integer maximus, felis nec lobortis vulputate, nunc leo blandit sem, id suscipit velit leo in purus.
+Pellentesque vel arcu sit amet nulla tincidunt ornare.
+Mauris sodales lectus eget est laoreet lobortis.
+
+\subsection[Reliablility]{Reliability}
+
+ Sed iaculis sodales elit quis vehicula.
+In et tristique neque, sodales aliquet metus.
+In posuere dictum nisl, quis laoreet augue congue a.
+Aenean in commodo neque, sit amet hendrerit ex.
+Aliquam id faucibus ante.
+Vivamus in fermentum nunc.
+Nam condimentum eros id orci pretium, quis aliquam magna eleifend.
+
+\section[Conclusions]{Conclusions}
+
 \end{document}

### 2015-08-03 inst/doc/Paper.Rnw
edits to methods section

@@
 
-Mauris ac lectus sagittis, mollis diam porttitor, efficitur purus.
-Duis ut eros vitae dolor vulputate sagittis.
-Aenean viverra, dui in interdum consectetur, lectus elit placerat urna, dictum ultricies ex justo eget est.
-In at leo quam.
-Suspendisse eros nulla, gravida nec suscipit vel, egestas eget neque.
-Etiam pharetra, nisl nec viverra ultricies, mauris lectus facilisis velit, vel aliquam ipsum lorem ut sem.
-Nulla ornare, justo at tempus blandit, nulla urna porta ligula, id dictum mi est ac massa.
-Quisque at lacus neque.
-Pellentesque est nulla, dignissim eget lobortis ut, vehicula at erat.
-Aenean ornare lacus mattis, elementum elit vel, tempor risus.
-In in purus tempor lacus imperdiet rhoncus nec nec tortor.
-Duis sagittis nisl ante, id egestas neque tristique fermentum.
-Fusce aliquet, odio non auctor aliquet, purus orci venenatis purus, sit amet pulvinar nisi est at dolor.
-Pellentesque lobortis dui eros, et ultricies tellus ultrices at.
-Ut sit amet interdum justo.
-Integer placerat vehicula interdum.
+To provide consistency we have copied all the data sets used in the
+timings to the \pkg{Timings} package itself. We have done all timings
+on the same computer. This computer has a relatively recent Intel
+processor and we used the
+\href{https://software.intel.com/en-us/intel-mkl}{Intel Math Kernel
+Library (MKL)} with \proglang{Julia}. We attempted to use
+\href{www.revolutionanalytics.com/revolution-r-open}{Revolution R Open
+(RRO)} as the \proglang{R} implementation as it can be configured with \pkg{MKL}.
+However, we ran into version problems with this so we used the standard
+Ubuntu version of \proglang{R} linked against OpenBLAS, which is also
+multi-threaded.
+
+Variables were renamed in the pattern:
+
+\begin{itemize}
+\item \textbf{Y} the response 
+\item \textbf{A}, \textbf{B}, \dots{} categorical covariates 
+\item \textbf{G},\textbf{H}, \textbf{I}, \dots{} grouping factors for random effects 
+\item \textbf{U}, \textbf{V}, \dots{} (skipping \textbf{Y}) continuous covariates
+\end{itemize}
+
+The timing results are saved in \href{http://json.org}{JSON (JavaScript
+Object Notation)} files in the directory accessible as
+
+<<eval=FALSE>>=
+system.file("JSON",package="Timings")
+@
+
+within \pkg{R}. The directory name will end with
+\texttt{./Timings/inst/JSON/} in the package source directory, for
+example the result of cloning the
+\href{https://github.com/Stat990-033/Timings}{github repository}. There
+is one \texttt{.json} file for each data set. Each such file contains
+results on timings of one or more models.
+
+The \pkg{Timings} package for \proglang{R} provides a \texttt{retime}
+function that takes the name of one of these JSON files and, optionally,
+the name of a file with the updated timings. Similarly there are some
+source files for Julia retimings.
+
+%    \begin{Verbatim}[commandchars=\\\{\}]
+%{\color{incolor}In [{\color{incolor}1}]:} \PY{n}{include}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../julia%%/retime.jl}\PY{l+s}{\PYZdq{}}\PY{p}{)}
+%        \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p%}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
+%       \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}%{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
+%\end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+dsname => "Alfalfa"
+    \end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+
+        `parse` has no method matching parse(::Int64)
+    while loading In[1], in expression starting on line 2
+
+         in retime at /home/bates/git/Timings/inst/julia/retime.jl:17
+
+    \end{Verbatim}
+
+    The timing was repeated so that compilation time is not included in the
+results. This repetition is only needed once per session.
+
+A careful examination of these results shows that the main differences
+in the Julia timings (the R timings are merely reported, not evaluated)
+are that the \texttt{LN\_BOBYQA} and \texttt{LD\_MMA} optimizers are
+much faster in the second run. This is because much of the code needs to
+be compiled the first time that a derivative-free optimizer and a
+derivative-based optimizer are used.
+
+The names of the optimizers used with \texttt{lmm} are those from the
+\href{https://github.com/JuliaOpt/NLopt.jl}{NLopt package} for
+\textbf{Julia}. Names that begin with \texttt{LD\_} are gradient-based
+methods. Names that begin with \texttt{LN\_} are derivative-free
+methods. There is one other derivative-free method, \texttt{LN\_PRAXIS},
+available in the \texttt{NLopt} package but, for some reason, it can
+hang on very simple problems like this. Frequently we omit it.
+
+The optimizers used with \texttt{lmer} include the \texttt{Nelder\_Mead}
+optimizer built into the \texttt{lme4} package, the \texttt{bobyqa}
+optimizer from the
+\href{http://cran.rstudio.com/web/packages/minqa/index.html}{minqa
+package}, the derivative-free optimizers from the
+\href{http://cran.rstudio.com/web/packages/nloptr/index.html}{nloptr
+package} and several optimizers from the
+\href{http://cran.rstudio.com/web/packages/optimx/index.html}{optimx
+pacakge}.
+
+The \texttt{optimx:bobyqa} optimizer is just a wrapper around
+\texttt{bobyqa} (bounded optimization by quadratic approximation) from
+the \texttt{minqa} package and should provide results similar to those
+from the \texttt{bobyqa} optimizer. For some reason the number of
+function evaluations is not reported for the version in \texttt{optimx}.
+
+The optimizers from \texttt{nloptr} (i.e.~those whose names begin with
+\texttt{NLOPT\_LN\_}) use the same underlying code as do the similarly
+named optimizers in the \texttt{NLopt} package for \textbf{Julia}. The
+number of iterations to convergence should be similar for the same
+underlying code, although not nessarily exactly the same because the
+evaluation of the objective in \textbf{R} and in \textbf{Julia} may
+produce slightly different answers. Also the convergence criteria in the
+\textbf{Julia} version are more strict than those in the \textbf{R}
+version
+
+Also shown are the value of the criterion (negative twice the
+log-likelihood, lower is better) achieved, the elapsed time and the
+number of function and gradient evaluations. The \texttt{nopt} value is
+the number of parameters in the optimization problem. \texttt{mtype} is
+the model type in the \textbf{Julia} code. There are special methods for
+solving the penalized least squares (PLS) problem, and for evaluating
+the objective and its gradient when there is only one grouping factor
+for the random effects. The model type is called \texttt{PLSOne}.
+
+The \texttt{Alfalfa} example is a particularly easy one and all of the
+optimizerws converge to an objective value close to -10.81023 in less
+than 0.6 seconds.
 

### 2015-08-03 inst/doc/Paper.tex
edits to methods section

@@
 
-Mauris ac lectus sagittis, mollis diam porttitor, efficitur purus.
-Duis ut eros vitae dolor vulputate sagittis.
-Aenean viverra, dui in interdum consectetur, lectus elit placerat urna, dictum ultricies ex justo eget est.
-In at leo quam.
-Suspendisse eros nulla, gravida nec suscipit vel, egestas eget neque.
-Etiam pharetra, nisl nec viverra ultricies, mauris lectus facilisis velit, vel aliquam ipsum lorem ut sem.
-Nulla ornare, justo at tempus blandit, nulla urna porta ligula, id dictum mi est ac massa.
-Quisque at lacus neque.
-Pellentesque est nulla, dignissim eget lobortis ut, vehicula at erat.
-Aenean ornare lacus mattis, elementum elit vel, tempor risus.
-In in purus tempor lacus imperdiet rhoncus nec nec tortor.
-Duis sagittis nisl ante, id egestas neque tristique fermentum.
-Fusce aliquet, odio non auctor aliquet, purus orci venenatis purus, sit amet pulvinar nisi est at dolor.
-Pellentesque lobortis dui eros, et ultricies tellus ultrices at.
-Ut sit amet interdum justo.
-Integer placerat vehicula interdum.
+To provide consistency we have copied all the data sets used in the
+timings to the \pkg{Timings} package itself. We have done all timings
+on the same computer. This computer has a relatively recent Intel
+processor and we used the
+\href{https://software.intel.com/en-us/intel-mkl}{Intel Math Kernel
+Library (MKL)} with \proglang{Julia}. We attempted to use
+\href{www.revolutionanalytics.com/revolution-r-open}{Revolution R Open
+(RRO)} as the \proglang{R} implementation as it can be configured with \pkg{MKL}.
+However, we ran into version problems with this so we used the standard
+Ubuntu version of R linked against OpenBLAS, which is also
+multi-threaded.
+
+Variables were renamed in the pattern:
+
+\begin{itemize}
+\item \textbf{Y} the response 
+\item \textbf{A}, \textbf{B}, \dots{} categorical covariates 
+\item \textbf{G},\textbf{H}, \textbf{I}, \dots{} grouping factors for random effects 
+\item \textbf{U}, \textbf{V}, \dots{} (skipping \textbf{Y}) continuous covariates
+\end{itemize}
+
+The timing results are saved in \href{http://json.org}{JSON (JavaScript
+Object Notation)} files in the directory accessible as
+
+\begin{Schunk}
+\begin{Sinput}
+> system.file("JSON",package="Timings")
+\end{Sinput}
+\end{Schunk}
+
+within \pkg{R}. The directory name will end with
+\texttt{./Timings/inst/JSON/} in the package source directory, for
+example the result of cloning the
+\href{https://github.com/Stat990-033/Timings}{github repository}. There
+is one \texttt{.json} file for each data set. Each such file contains
+results on timings of one or more models.
+
+The \pkg{Timings} package for \proglang{R} provides a \texttt{retime}
+function that takes the name of one of these JSON files and, optionally,
+the name of a file with the updated timings. Similarly there are some
+source files for Julia retimings.
+
+%    \begin{Verbatim}[commandchars=\\\{\}]
+%{\color{incolor}In [{\color{incolor}1}]:} \PY{n}{include}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../julia%%/retime.jl}\PY{l+s}{\PYZdq{}}\PY{p}{)}
+%        \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p%}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
+%       \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}%{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
+%\end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+dsname => "Alfalfa"
+    \end{Verbatim}
+
+    \begin{Verbatim}[commandchars=\\\{\}]
+
+        `parse` has no method matching parse(::Int64)
+    while loading In[1], in expression starting on line 2
+
+         in retime at /home/bates/git/Timings/inst/julia/retime.jl:17
+
+    \end{Verbatim}
+
+    The timing was repeated so that compilation time is not included in the
+results. This repetition is only needed once per session.
+
+A careful examination of these results shows that the main differences
+in the Julia timings (the R timings are merely reported, not evaluated)
+are that the \texttt{LN\_BOBYQA} and \texttt{LD\_MMA} optimizers are
+much faster in the second run. This is because much of the code needs to
+be compiled the first time that a derivative-free optimizer and a
+derivative-based optimizer are used.
+
+The names of the optimizers used with \texttt{lmm} are those from the
+\href{https://github.com/JuliaOpt/NLopt.jl}{NLopt package} for
+\textbf{Julia}. Names that begin with \texttt{LD\_} are gradient-based
+methods. Names that begin with \texttt{LN\_} are derivative-free
+methods. There is one other derivative-free method, \texttt{LN\_PRAXIS},
+available in the \texttt{NLopt} package but, for some reason, it can
+hang on very simple problems like this. Frequently we omit it.
+
+The optimizers used with \texttt{lmer} include the \texttt{Nelder\_Mead}
+optimizer built into the \texttt{lme4} package, the \texttt{bobyqa}
+optimizer from the
+\href{http://cran.rstudio.com/web/packages/minqa/index.html}{minqa
+package}, the derivative-free optimizers from the
+\href{http://cran.rstudio.com/web/packages/nloptr/index.html}{nloptr
+package} and several optimizers from the
+\href{http://cran.rstudio.com/web/packages/optimx/index.html}{optimx
+pacakge}.
+
+The \texttt{optimx:bobyqa} optimizer is just a wrapper around
+\texttt{bobyqa} (bounded optimization by quadratic approximation) from
+the \texttt{minqa} package and should provide results similar to those
+from the \texttt{bobyqa} optimizer. For some reason the number of
+function evaluations is not reported for the version in \texttt{optimx}.
+
+The optimizers from \texttt{nloptr} (i.e.~those whose names begin with
+\texttt{NLOPT\_LN\_}) use the same underlying code as do the similarly
+named optimizers in the \texttt{NLopt} package for \textbf{Julia}. The
+number of iterations to convergence should be similar for the same
+underlying code, although not nessarily exactly the same because the
+evaluation of the objective in \textbf{R} and in \textbf{Julia} may
+produce slightly different answers. Also the convergence criteria in the
+\textbf{Julia} version are more strict than those in the \textbf{R}
+version
+
+Also shown are the value of the criterion (negative twice the
+log-likelihood, lower is better) achieved, the elapsed time and the
+number of function and gradient evaluations. The \texttt{nopt} value is
+the number of parameters in the optimization problem. \texttt{mtype} is
+the model type in the \textbf{Julia} code. There are special methods for
+solving the penalized least squares (PLS) problem, and for evaluating
+the objective and its gradient when there is only one grouping factor
+for the random effects. The model type is called \texttt{PLSOne}.
+
+The \texttt{Alfalfa} example is a particularly easy one and all of the
+optimizerws converge to an objective value close to -10.81023 in less
+than 0.6 seconds.
 

### 2015-08-03 inst/doc/Paper.Rnw
more work on methods, work on results, edits to latex tables

@@
 the name of a file with the updated timings. Similarly there are some
-source files for Julia retimings.
+source files for \pkg{Julia} retimings.
 
-%    \begin{Verbatim}[commandchars=\\\{\}]
-%{\color{incolor}In [{\color{incolor}1}]:} \PY{n}{include}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../julia%%/retime.jl}\PY{l+s}{\PYZdq{}}\PY{p}{)}
-%        \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p%}{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
-%       \PY{n}{retime}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{../JSON/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}%{,}\PY{l+s}{\PYZdq{}}\PY{l+s}{/tmp/Alfalfa.json}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
-%\end{Verbatim}
+<<eval=FALSE>>=
+include("../julia/retime.jl")
+retime("../JSON/Alfalfa.json","/tmp/Alfalfa.json")
+retime("../JSON/Alfalfa.json","/tmp/Alfalfa.json")
+@
 
-    \begin{Verbatim}[commandchars=\\\{\}]
-dsname => "Alfalfa"
-    \end{Verbatim}
-
-    \begin{Verbatim}[commandchars=\\\{\}]
-
-        `parse` has no method matching parse(::Int64)
-    while loading In[1], in expression starting on line 2
-
-         in retime at /home/bates/git/Timings/inst/julia/retime.jl:17
-
-    \end{Verbatim}
+INCLUDE SUMMARY TABLE USING res.rda
 
@@
 A careful examination of these results shows that the main differences
-in the Julia timings (the R timings are merely reported, not evaluated)
+in the \pkg{Julia} timings (the \pkg{R} timings are merely reported, not evaluated)
 are that the \texttt{LN\_BOBYQA} and \texttt{LD\_MMA} optimizers are
@@
 The names of the optimizers used with \texttt{lmm} are those from the
-\href{https://github.com/JuliaOpt/NLopt.jl}{NLopt package} for
-\textbf{Julia}. Names that begin with \texttt{LD\_} are gradient-based
+\pkg{NLopt} package for \proglang{Julia}. Names that begin with \texttt{LD\_} are gradient-based
 methods. Names that begin with \texttt{LN\_} are derivative-free
 methods. There is one other derivative-free method, \texttt{LN\_PRAXIS},
-available in the \texttt{NLopt} package but, for some reason, it can
+available in the \pkg{NLopt} package but, for some reason, it can
 hang on very simple problems like this. Frequently we omit it.
@@
 The optimizers used with \texttt{lmer} include the \texttt{Nelder\_Mead}
-optimizer built into the \texttt{lme4} package, the \texttt{bobyqa}
+optimizer built into the \pkg{lme4} package, the \texttt{bobyqa}
 optimizer from the
-\href{http://cran.rstudio.com/web/packages/minqa/index.html}{minqa
-package}, the derivative-free optimizers from the
-\href{http://cran.rstudio.com/web/packages/nloptr/index.html}{nloptr
-package} and several optimizers from the
-\href{http://cran.rstudio.com/web/packages/optimx/index.html}{optimx
-pacakge}.
+\pkg{minqa} package, the derivative-free optimizers from the
+\pkg{nloptr} package and several optimizers from the
+\pkg{optimx} pacakge.
 
@@
 \texttt{bobyqa} (bounded optimization by quadratic approximation) from
-the \texttt{minqa} package and should provide results similar to those
+the \pkg{minqa} package and should provide results similar to those
 from the \texttt{bobyqa} optimizer. For some reason the number of
-function evaluations is not reported for the version in \texttt{optimx}.
+function evaluations is not reported for the version in \pkg{optimx}.
 
-The optimizers from \texttt{nloptr} (i.e.~those whose names begin with
+The optimizers from \pkg{nloptr} (i.e.~those whose names begin with
 \texttt{NLOPT\_LN\_}) use the same underlying code as do the similarly
-named optimizers in the \texttt{NLopt} package for \textbf{Julia}. The
+named optimizers in the \pkg{NLopt} package for \proglang{Julia}. The
 number of iterations to convergence should be similar for the same
 underlying code, although not nessarily exactly the same because the
-evaluation of the objective in \textbf{R} and in \textbf{Julia} may
+evaluation of the objective in \proglang{R} and in \proglang{Julia} may
 produce slightly different answers. Also the convergence criteria in the
-\textbf{Julia} version are more strict than those in the \textbf{R}
+\proglang{Julia} version are more strict than those in the \proglang{R}
 version
@@
 the number of parameters in the optimization problem. \texttt{mtype} is
-the model type in the \textbf{Julia} code. There are special methods for
+the model type in the \proglang{Julia} code. There are special methods for
 solving the penalized least squares (PLS) problem, and for evaluating
@@
 
-Nullam eget gravida justo.
-Proin lectus mauris, dictum vitae ultricies ut, accumsan sed nisi.
-Pellentesque quis nunc vestibulum, aliquam arcu eu, dapibus risus.
-Pellentesque lobortis nulla at posuere molestie.
-Etiam risus nisi, pellentesque a tortor at, porta mollis est.
-Curabitur nec ante eget felis pellentesque sagittis id in nulla.
-Vivamus sit amet accumsan lorem.
-Nam egestas ultrices nisi, volutpat rutrum tortor tincidunt ac.
-Ut ex nunc, ornare eget nibh quis, mollis aliquam dui.
-Sed tempor, sapien sed vestibulum malesuada, dolor est ullamcorper leo, at mollis erat nulla eget urna.
-Aenean risus leo, pellentesque sit amet pulvinar a, ultrices ut lacus.
-Sed id nunc tempus, commodo augue a, hendrerit nunc.
-Mauris et purus lorem.
-Praesent ornare, ipsum id bibendum placerat, arcu enim iaculis tellus, tincidunt rhoncus augue dui sed ex.
-Proin nunc nisl, lacinia quis tristique eget, tempor eu velit.
+For the \texttt{Alfalfa} data there is not much of a burden in refitting
+the model with all the \textbf{Julia} optimizers just to get the table
+shown above. But other examples can take an hour or more to converge and
+we don't really need to refit them every time. The \texttt{tabulate.jl}
+file contains a function \texttt{optdir} to create a \texttt{DataFrame}
+from the results of all the model fits.
 
-\subsection[Speed]{Speed}
+<<eval=FALSE>>=
+include("../julia/tabulate.jl")
+res = optdir("../JSON");
+
+@
+
+\begin{verbatim}
+  | Row | opt        | dsname          | n      | np | excess  | time    | reltime|    
+  |-----|------------|-----------------|--------|----|---------|---------|-------|
+  | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  | 27.5171|
+  | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 13.9236|
+  | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 10.8009|
+  | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  | 11.7005|
+  | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  | 13.6969|
+  | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  | 12.608|
+  | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  | 23.2699|
+  | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  | 18.8753|
+  | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  | 15.6477|
+  | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  | 22.0827|
+  | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
+  .
+  .
+  .
+  | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
+  | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  | 8.6481|
+  | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  | 3.7438|
+  | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   | 5.0422|
+  | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 | 9.2732|
+  | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  | 15.9582|
+  | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  | 1.1034|
+  | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  | 5.3039|
+  | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  | 8.6177|
+  | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  | 15.3763|
+  | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  | 1.133|
+  | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   | 2.3154|
+
+\end{verbatim}
+
+\subsection[Convergence]{Proportion Converged}
  Etiam placerat commodo dapibus.
@@
 
+\section[References]{References}
+
+Note all optimization packages used, lmm, lme4, jsonlite,
+
 \end{document}

### 2015-08-03 inst/doc/Paper.tex
more work on methods, work on results, edits to latex tables

--- 2015-07-31 inst/doc/Paper.tex
@@
 
-Mauris ac lectus sagittis, mollis diam porttitor, efficitur purus.
-Duis ut eros vitae dolor vulputate sagittis.
-Aenean viverra, dui in interdum consectetur, lectus elit placerat urna, dictum ultricies ex justo eget est.
-In at leo quam.
-Suspendisse eros nulla, gravida nec suscipit vel, egestas eget neque.
-Etiam pharetra, nisl nec viverra ultricies, mauris lectus facilisis velit, vel aliquam ipsum lorem ut sem.
-Nulla ornare, justo at tempus blandit, nulla urna porta ligula, id dictum mi est ac massa.
-Quisque at lacus neque.
-Pellentesque est nulla, dignissim eget lobortis ut, vehicula at erat.
-Aenean ornare lacus mattis, elementum elit vel, tempor risus.
-In in purus tempor lacus imperdiet rhoncus nec nec tortor.
-Duis sagittis nisl ante, id egestas neque tristique fermentum.
-Fusce aliquet, odio non auctor aliquet, purus orci venenatis purus, sit amet pulvinar nisi est at dolor.
-Pellentesque lobortis dui eros, et ultricies tellus ultrices at.
-Ut sit amet interdum justo.
-Integer placerat vehicula interdum.
+To provide consistency we have copied all the data sets used in the
+timings to the \pkg{Timings} package itself. We have done all timings
+on the same computer. This computer has a relatively recent Intel
+processor and we used the
+\href{https://software.intel.com/en-us/intel-mkl}{Intel Math Kernel
+Library (MKL)} with \proglang{Julia}. We attempted to use
+\href{www.revolutionanalytics.com/revolution-r-open}{Revolution R Open
+(RRO)} as the \proglang{R} implementation as it can be configured with \pkg{MKL}.
+However, we ran into version problems with this so we used the standard
+Ubuntu version of \proglang{R} linked against OpenBLAS, which is also
+multi-threaded.
+
+Variables were renamed in the pattern:
+
+\begin{itemize}
+\item \textbf{Y} the response 
+\item \textbf{A}, \textbf{B}, \dots{} categorical covariates 
+\item \textbf{G},\textbf{H}, \textbf{I}, \dots{} grouping factors for random effects 
+\item \textbf{U}, \textbf{V}, \dots{} (skipping \textbf{Y}) continuous covariates
+\end{itemize}
+
+The timing results are saved in \href{http://json.org}{JSON (JavaScript
+Object Notation)} files in the directory accessible as
+
+\begin{Schunk}
+\begin{Sinput}
+> system.file("JSON",package="Timings")
+\end{Sinput}
+\end{Schunk}
+
+within \pkg{R}. The directory name will end with
+\texttt{./Timings/inst/JSON/} in the package source directory, for
+example the result of cloning the
+\href{https://github.com/Stat990-033/Timings}{github repository}. There
+is one \texttt{.json} file for each data set. Each such file contains
+results on timings of one or more models.
+
+The \pkg{Timings} package for \proglang{R} provides a \texttt{retime}
+function that takes the name of one of these JSON files and, optionally,
+the name of a file with the updated timings. Similarly there are some
+source files for \pkg{Julia} retimings.
+
+\begin{Schunk}
+\begin{Sinput}
+> include("../julia/retime.jl")
+> retime("../JSON/Alfalfa.json","/tmp/Alfalfa.json")
+> retime("../JSON/Alfalfa.json","/tmp/Alfalfa.json")
+\end{Sinput}
+\end{Schunk}
+
+INCLUDE SUMMARY TABLE USING res.rda
+
+  The timing was repeated so that compilation time is not included in the
+results. This repetition is only needed once per session.
+
+A careful examination of these results shows that the main differences
+in the \pkg{Julia} timings (the \pkg{R} timings are merely reported, not evaluated)
+are that the \texttt{LN\_BOBYQA} and \texttt{LD\_MMA} optimizers are
+much faster in the second run. This is because much of the code needs to
+be compiled the first time that a derivative-free optimizer and a
+derivative-based optimizer are used.
+
+The names of the optimizers used with \texttt{lmm} are those from the
+\pkg{NLopt} package for \proglang{Julia}. Names that begin with \texttt{LD\_} are gradient-based
+methods. Names that begin with \texttt{LN\_} are derivative-free
+methods. There is one other derivative-free method, \texttt{LN\_PRAXIS},
+available in the \pkg{NLopt} package but, for some reason, it can
+hang on very simple problems like this. Frequently we omit it.
+
+The optimizers used with \texttt{lmer} include the \texttt{Nelder\_Mead}
+optimizer built into the \pkg{lme4} package, the \texttt{bobyqa}
+optimizer from the
+\pkg{minqa} package, the derivative-free optimizers from the
+\pkg{nloptr} package and several optimizers from the
+\pkg{optimx} pacakge.
+
+The \texttt{optimx:bobyqa} optimizer is just a wrapper around
+\texttt{bobyqa} (bounded optimization by quadratic approximation) from
+the \pkg{minqa} package and should provide results similar to those
+from the \texttt{bobyqa} optimizer. For some reason the number of
+function evaluations is not reported for the version in \pkg{optimx}.
+
+The optimizers from \pkg{nloptr} (i.e.~those whose names begin with
+\texttt{NLOPT\_LN\_}) use the same underlying code as do the similarly
+named optimizers in the \pkg{NLopt} package for \proglang{Julia}. The
+number of iterations to convergence should be similar for the same
+underlying code, although not nessarily exactly the same because the
+evaluation of the objective in \proglang{R} and in \proglang{Julia} may
+produce slightly different answers. Also the convergence criteria in the
+\proglang{Julia} version are more strict than those in the \proglang{R}
+version
+
+Also shown are the value of the criterion (negative twice the
+log-likelihood, lower is better) achieved, the elapsed time and the
+number of function and gradient evaluations. The \texttt{nopt} value is
+the number of parameters in the optimization problem. \texttt{mtype} is
+the model type in the \proglang{Julia} code. There are special methods for
+solving the penalized least squares (PLS) problem, and for evaluating
+the objective and its gradient when there is only one grouping factor
+for the random effects. The model type is called \texttt{PLSOne}.
+
+The \texttt{Alfalfa} example is a particularly easy one and all of the
+optimizerws converge to an objective value close to -10.81023 in less
+than 0.6 seconds.
 
@@
 
-Nullam eget gravida justo.
-Proin lectus mauris, dictum vitae ultricies ut, accumsan sed nisi.
-Pellentesque quis nunc vestibulum, aliquam arcu eu, dapibus risus.
-Pellentesque lobortis nulla at posuere molestie.
-Etiam risus nisi, pellentesque a tortor at, porta mollis est.
-Curabitur nec ante eget felis pellentesque sagittis id in nulla.
-Vivamus sit amet accumsan lorem.
-Nam egestas ultrices nisi, volutpat rutrum tortor tincidunt ac.
-Ut ex nunc, ornare eget nibh quis, mollis aliquam dui.
-Sed tempor, sapien sed vestibulum malesuada, dolor est ullamcorper leo, at mollis erat nulla eget urna.
-Aenean risus leo, pellentesque sit amet pulvinar a, ultrices ut lacus.
-Sed id nunc tempus, commodo augue a, hendrerit nunc.
-Mauris et purus lorem.
-Praesent ornare, ipsum id bibendum placerat, arcu enim iaculis tellus, tincidunt rhoncus augue dui sed ex.
-Proin nunc nisl, lacinia quis tristique eget, tempor eu velit.
+For the \texttt{Alfalfa} data there is not much of a burden in refitting
+the model with all the \textbf{Julia} optimizers just to get the table
+shown above. But other examples can take an hour or more to converge and
+we don't really need to refit them every time. The \texttt{tabulate.jl}
+file contains a function \texttt{optdir} to create a \texttt{DataFrame}
+from the results of all the model fits.
 
-\subsection[Speed]{Speed}
+\begin{Schunk}
+\begin{Sinput}
+> include("../julia/tabulate.jl")
+> res = optdir("../JSON");
+> 
+\end{Sinput}
+\end{Schunk}
+
+\begin{verbatim}
+  | Row | opt        | dsname          | n      | np | excess  | time    | reltime|    
+  |-----|------------|-----------------|--------|----|---------|---------|-------|
+  | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  | 27.5171|
+  | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 13.9236|
+  | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 10.8009|
+  | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  | 11.7005|
+  | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  | 13.6969|
+  | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  | 12.608|
+  | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  | 23.2699|
+  | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  | 18.8753|
+  | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  | 15.6477|
+  | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  | 22.0827|
+  | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
+  .
+  .
+  .
+  | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
+  | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  | 8.6481|
+  | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  | 3.7438|
+  | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   | 5.0422|
+  | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 | 9.2732|
+  | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  | 15.9582|
+  | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  | 1.1034|
+  | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  | 5.3039|
+  | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  | 8.6177|
+  | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  | 15.3763|
+  | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  | 1.133|
+  | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   | 2.3154|
+
+\end{verbatim}
+
+\subsection[Convergence]{Proportion Converged}
  Etiam placerat commodo dapibus.

### 2015-08-03 inst/doc/Paper.Rnw
Edited results section, tried to fix issues with latex table.  Note that pdf from rstuio will not directly compile as written, but pdf file is still created correctly.

@@
   | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
-  .
-  .
-  .
+  | 12  | "LD\_CCSAQ" | "Dyestuff2"     | 30     | 1  | 0.0     | 0.0005  | 0.682|
+  | 13  | "LD\_CCSAQ" | "Dyestuff"      | 30     | 1  | 0.0     | 0.0008  | 1.0184|
+  | 14  | "LD\_CCSAQ" | "ergoStool"     | 36     | 1  | 0.0     | 0.0012  | 1.1018|
+  | 15  | "LD\_CCSAQ" | "Exam"          | 4059   | 1  | 0.0     | 0.0105  | 1.1851|
+  | 16  | "LD\_CCSAQ" | "Exam"          | 4059   | 1  | 0.0     | 0.0105  | 1.318|
+  | 17  | "LD\_CCSAQ" | "Gasoline"      | 32     | 1  | 0.0     | 0.0011  | 0.8193|
+  | 18  | "LD\_CCSAQ" | "gb12"          | 225866 | 6  | 0.0     | 3.5221  | 17.4791|
   | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
@@
 
+\end{verbatim}
+
+The \texttt{time} column is the time in seconds to converge. The
+\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
+optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+For \proglang{Julia} the time column is the time in seconds to converge. The reltime column is the time relative to the LN_BOBYQA optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+%%% Note that because of this table, rstudio cannot compile pdf because of margins,
+% but pdf is still available (and compiled properly) as seperate file.
+\begin{verbatim}
+| Row | opt               | dsname          | n      | np | excess | time   | reltime|
+|-----|-------------------|-----------------|--------|----|--------|--------| ----|
+| 1|"NLOPT\_LN\_BOBYQA" | "Alfalfa"      | 72     | 1  | 0.0    | 0.042  | 27.5171|
+| 2|"NLOPT\_LN\_BOBYQA" | "Animal"       | 20     | 2  | 0.0    | 0.023  |13.9236|
+| 3|"NLOPT\_LN\_BOBYQA" | "Assay"        | 60     | 2  | 1.0e-5 | 0.032  |10.8009|
+| 4|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |11.7005|
+| 5|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |13.6969|
+| 6|"NLOPT\_LN\_BOBYQA" | "BIB"          | 24     | 1  | 0.0    | 0.02   |12.608|
+| 7|"NLOPT\_LN\_BOBYQA" | "Bond"         | 21     | 1  | 0.0    | 0.02   |23.2699
+| 8|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 20 | 1.0e-5 | 4.661  |18.8753|
+| 9|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 8  | 0.0    | 1.057  |15.6477|
+| 10|"NLOPT\_LN\_BOBYQA" | "cake"        | 270    | 1  | 0.0    | 0.053  |22.0827
+| 11|"NLOPT\_LN\_BOBYQA" | "Chem97"      | 31022  | 2  | 0.0    | 0.632  |3.942|
+.
+.
+.
+| 38|"NLOPT\_LN\_BOBYQA" | "PBIB"        | 60     | 1  | 0.0    | 0.018  |13.1358|
+| 39|"NLOPT\_LN\_BOBYQA" | "Penicillin"  | 144    | 2  | 0.0    | 0.023  |8.6481|
+| 40|"NLOPT\_LN\_BOBYQA" | "Poems"       | 275996 | 3  | 0.0    | 21.309 |3.7438|
+| 41|"NLOPT\_LN\_BOBYQA" | "ScotsSec"    | 3435   | 2  | 0.0    | 0.076  |5.0422|
+| 42|"NLOPT\_LN\_BOBYQA" | "Semi2"       | 72     | 3  | 0.0    | 0.03   |9.2732|
+| 43|"NLOPT\_LN\_BOBYQA" | "Semiconductor"| 48     | 1  | 0.0    | 0.019  |15.9582|
+| 44|"NLOPT\_LN\_BOBYQA" | "SIMS"        | 3691   | 3  | 0.0    | 0.15   |1.1034|
+| 45|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 3  | 0.0    | 0.037  |5.3039|
+| 46|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 2  | 0.0    | 0.024  |15.3763|
+| 47|"NLOPT\_LN\_BOBYQA" | "TeachingII"  | 96     | 1  | 0.0    | 0.021  |1.133|
+| 48|"NLOPT\_LN\_BOBYQA" | "Weights"     | 399    | 3  | 1.0e-5 | 0.039  |5.3039|
+| 49|"NLOPT\_LN\_BOBYQA" | "WWheat"      | 60     | 3  | 0.0    | 0.025  |2.3154|
 \end{verbatim}

### 2015-08-03 inst/doc/Paper.tex
Edited results section, tried to fix issues with latex table.  Note that pdf from rstuio will not directly compile as written, but pdf file is still created correctly.

@@
 
+\end{verbatim}
+
+The \texttt{time} column is the time in seconds to converge. The
+\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
+optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+For \proglang{Julia} the time column is the time in seconds to converge. The reltime column is the time relative to the LN_BOBYQA optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+\begin{verbatim}
+| Row | opt               | dsname          | n      | np | excess | time   | reltime|
+|-----|-------------------|-----------------|--------|----|--------|--------| ----|
+| 1|"NLOPT\_LN\_BOBYQA" | "Alfalfa"      | 72     | 1  | 0.0    | 0.042  | 27.5171|
+| 2|"NLOPT\_LN\_BOBYQA" | "Animal"       | 20     | 2  | 0.0    | 0.023  |13.9236|
+| 3|"NLOPT\_LN\_BOBYQA" | "Assay"        | 60     | 2  | 1.0e-5 | 0.032  |10.8009|
+| 4|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |11.7005|
+| 5|"NLOPT\_LN\_BOBYQA" | "AvgDailyGain" | 32     | 1  | 0.0    | 0.02   |13.6969|
+| 6|"NLOPT\_LN\_BOBYQA" | "BIB"          | 24     | 1  | 0.0    | 0.02   |12.608|
+| 7|"NLOPT\_LN\_BOBYQA" | "Bond"         | 21     | 1  | 0.0    | 0.02   |23.2699
+| 8|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 20 | 1.0e-5 | 4.661  |18.8753|
+| 9|"NLOPT\_LN\_BOBYQA" | "bs10"         | 1104   | 8  | 0.0    | 1.057  |15.6477|
+| 10|"NLOPT\_LN\_BOBYQA" | "cake"        | 270    | 1  | 0.0    | 0.053  |22.0827
+| 11|"NLOPT\_LN\_BOBYQA" | "Chem97"      | 31022  | 2  | 0.0    | 0.632  |3.942|
+.
+.
+.
+| 38|"NLOPT\_LN\_BOBYQA" | "PBIB"        | 60     | 1  | 0.0    | 0.018  |13.1358|
+| 39|"NLOPT\_LN\_BOBYQA" | "Penicillin"  | 144    | 2  | 0.0    | 0.023  |8.6481|
+| 40|"NLOPT\_LN\_BOBYQA" | "Poems"       | 275996 | 3  | 0.0    | 21.309 |3.7438|
+| 41|"NLOPT\_LN\_BOBYQA" | "ScotsSec"    | 3435   | 2  | 0.0    | 0.076  |5.0422|
+| 42|"NLOPT\_LN\_BOBYQA" | "Semi2"       | 72     | 3  | 0.0    | 0.03   |9.2732|
+| 43|"NLOPT\_LN\_BOBYQA" | "Semiconductor"| 48     | 1  | 0.0    | 0.019  |15.9582|
+| 44|"NLOPT\_LN\_BOBYQA" | "SIMS"        | 3691   | 3  | 0.0    | 0.15   |1.1034|
+| 45|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 3  | 0.0    | 0.037  |5.3039|
+| 46|"NLOPT\_LN\_BOBYQA" | "sleepstudy"  | 180    | 2  | 0.0    | 0.024  |15.3763|
+| 47|"NLOPT\_LN\_BOBYQA" | "TeachingII"  | 96     | 1  | 0.0    | 0.021  |1.133|
+| 48|"NLOPT\_LN\_BOBYQA" | "Weights"     | 399    | 3  | 1.0e-5 | 0.039  |5.3039|
+| 49|"NLOPT\_LN\_BOBYQA" | "WWheat"      | 60     | 3  | 0.0    | 0.025  |2.3154|
 \end{verbatim}

### 2015-08-04 inst/doc/Paper.Rnw
Filled in later sections for a very, very rough outline.  NOTE .Rnw will still not compile on Rstudio, but pdfis still created correctly

@@
 \subsection[Convergence]{Proportion Converged}
- Etiam placerat commodo dapibus.
-Vivamus a porta dui.
-Phasellus accumsan sollicitudin cursus.
-In nec lorem sit amet purus vulputate suscipit quis ac lectus.
-Aliquam sollicitudin malesuada blandit.
-Aliquam sed nisi accumsan, pulvinar diam eget, facilisis urna.
-Fusce a dignissim ipsum.
-Integer maximus, felis nec lobortis vulputate, nunc leo blandit sem, id suscipit velit leo in purus.
-Pellentesque vel arcu sit amet nulla tincidunt ornare.
-Mauris sodales lectus eget est laoreet lobortis.
+
+The most important question regarding the optimizers is whether or not they have converged to the global optimum.
+We cannot test this directly.
+Instead we use a "crowd-sourced" criterion based on the minimum objective achieved by any of the algorithms.
+The difference between the objective achieved by a particular algorithm and this minimum is called the excess.
+In the summaries \texttt{excess} is rounded to 5 digits after the decimal so the minimum non-zero \texttt{excess} is $10^{-5}$.
+
+ \begin{Verbatim}[commandchars=\\\{\}]
+{\color{outcolor}Out[{\color{outcolor}4}]:} 49x3 DataFrame
+        | Row | opt         | dsname          | excess |
+        |-----|-------------|-----------------|--------|
+        | 1   | "LN\_BOBYQA" | "Alfalfa"       | 0.0    |
+        | 2   | "LN\_BOBYQA" | "Animal"        | 0.0    |
+        | 3   | "LN\_BOBYQA" | "Assay"         | 0.0    |
+        | 4   | "LN\_BOBYQA" | "AvgDailyGain"  | 0.0    |
+        | 5   | "LN\_BOBYQA" | "AvgDailyGain"  | 0.0    |
+        | 6   | "LN\_BOBYQA" | "BIB"           | 0.0    |
+        | 7   | "LN\_BOBYQA" | "Bond"          | 0.0    |
+        | 8   | "LN\_BOBYQA" | "bs10"          | 1.0e-5 |
+        | 9   | "LN\_BOBYQA" | "bs10"          | 0.0    |
+        | 10  | "LN\_BOBYQA" | "cake"          | 0.0    |
+        | 11  | "LN\_BOBYQA" | "Chem97"        | 0.0    |
+        \end{verbatim}
+        \vdots
+        \begin{verbatim}
+        | 38  | "LN\_BOBYQA" | "PBIB"          | 0.0    |
+        | 39  | "LN\_BOBYQA" | "Penicillin"    | 0.0    |
+        | 40  | "LN\_BOBYQA" | "Poems"         | 0.0    |
+        | 41  | "LN\_BOBYQA" | "ScotsSec"      | 0.0    |
+        | 42  | "LN\_BOBYQA" | "Semi2"         | 0.0    |
+        | 43  | "LN\_BOBYQA" | "Semiconductor" | 0.0    |
+        | 44  | "LN\_BOBYQA" | "SIMS"          | 0.0    |
+        | 45  | "LN\_BOBYQA" | "sleepstudy"    | 0.0    |
+        | 46  | "LN\_BOBYQA" | "sleepstudy"    | 0.0    |
+        | 47  | "LN\_BOBYQA" | "TeachingII"    | 0.0    |
+        | 48  | "LN\_BOBYQA" | "Weights"       | 0.0    |
+        | 49  | "LN\_BOBYQA" | "WWheat"        | 0.0    |
+\end{Verbatim}
+
+  If we wish to declare ``converged'' or ``not converged'' according to
+the excess objective value we must establish a threshold. An absolute
+threshold seems reasonable because the objective, negative twice the
+log-likelihood, is on a scale where differences in this objective are
+compared to a $\chi^2$ random variable. Thus an excess of $10^{-9}$ or
+even $10^{-5}$ is negligible.
+
+For each optimizer we can examine which of the data set/model
+combinations resulted in an excess greater than a threshold.
+
+            \begin{Verbatim}
+        | Row | opt                          | attempted | failed |
+        |-----|------------------------------|-----------|--------|
+        | 1   | "LD\_CCSAQ"                   | 35        | 11     |
+        | 2   | "LD\_LBFGS"                   | 35        | 11     |
+        | 3   | "LD\_MMA"                     | 36        | 5      |
+        | 4   | "LD\_SLSQP"                   | 35        | 4      |
+        | 5   | "LD\_TNEWTON"                 | 34        | 8      |
+        | 6   | "LD\_TNEWTON\_PRECOND"         | 34        | 8      |
+        | 7   | "LD\_TNEWTON\_PRECOND\_RESTART" | 33        | 12     |
+        | 8   | "LD\_TNEWTON\_RESTART"         | 34        | 11     |
+        | 9   | "LD\_VAR1"                    | 35        | 10     |
+        | 10  | "LD\_VAR2"                    | 35        | 10     |
+        | 11  | "LN\_BOBYQA"                  | 49        | 0      |
+        ⋮
+        | 15  | "LN\_SBPLX"                   | 49        | 2      |
+        | 16  | "NLOPT\_LN\_BOBYQA"            | 49        | 0      |
+        | 17  | "NLOPT\_LN\_COBYLA"            | 48        | 2      |
+        | 18  | "NLOPT\_LN\_NELDERMEAD"        | 47        | 6      |
+        | 19  | "NLOPT\_LN\_PRAXIS"            | 18        | 4      |
+        | 20  | "NLOPT\_LN\_SBPLX"             | 48        | 2      |
+        | 21  | "Nelder\_Mead"                | 49        | 8      |
+        | 22  | "bobyqa"                     | 49        | 2      |
+        | 23  | "optimx:L-BFGS-B"            | 49        | 0      |
+        | 24  | "optimx:bobyqa"              | 49        | 2      |
+        | 25  | "optimx:nlminb"              | 49        | 0      |
+        | 26  | "optimx:spg"                 | 49        | 4      |
+\end{verbatim}
+
+    At this threshold the most reliable algorithm in Julia is
+\texttt{LN\_BOBYQA}. In \proglang{R} the most reliable algorithms are
+\texttt{NLOPT\_LN\_BOBYQA}, \texttt{optimx:L-BFGS-B} and
+\texttt{optimx:nlminb}. It is interesting that nlminb is reliable as I
+felt that it wasn't converging well when it was the default optimizer in
+\pkg{lmer}.
+
+Interestingly, the derviative-based algorithms in \texttt{NLopt} were
+not as reliable as the derivative-free algorithms. The most likely
+explanation is that I don't have the gradient coded properly.
+
+The Nelder-Mead simplex algorithm did not perform well, failing on 8 out
+of 48 cases. For many of these the value at which convergence was
+declared was far from the optimum.
+
+\begin{verbatim}
+| Row | opt           | dsname        | excess  | time    | reltime | np | n|
+|-----|---------------|---------------|---------|---------|---------|----|--|
+| 1   | "Nelder\_Mead" | "bs10"        | 71.3859 | 145.368 | 588.684 | 20 |1104|
+| 2   | "Nelder\_Mead" | "d3"          | 317.59  | 454.519 | 4.2502  | 9  |130418|
+| 3   | "Nelder\_Mead" | "dialectNL"   | 181.632 | 54.541  | 31.1594 | 6  |225866|
+| 4   | "Nelder\_Mead" | "gb12"        | 78.7119 | 38.206  | 189.605 | 20 |512|
+| 5   | "Nelder\_Mead" | "kb07"        | 398.732 | 2825.46 | 667.015 | 72 |1790|
+| 6   | "Nelder\_Mead" | "kb07"        | 403.478 | 269.436 | 383.087 | 16 |1790|
+| 7   | "Nelder\_Mead" | "Mississippi" | 0.04272 | 0.018   | 20.3989 | 1  |37|
+| 8   | "Nelder\_Mead" | "mm0"         | 181.632 | 76.87   | 70.0716 | 6  |69588|  
+\end{verbatim}
+
+ The \texttt{Nelder\_Mead} algorithm, either in the native form in
+\pkg{lmer} or in the \texttt{NLopt} implementation performed poorly
+on those cases with many parameters to optimize. It was both unreliable
+and slow, taking over 45 minutes to reach a spurious optimum on the
+``maximal'' model (in the sense of Barr et al., 2012) for the
+\texttt{kb07} data from Kronmueller and Barr (2007). This is not
+terribly surprising given that the model is horribly overparameterized,
+but still it shows that this algorithm is not a good choice in these
+cases.
+
+We note in passing that all the models involving fitting 20 or more
+parameters are ``maximal'' models in the sense of Barr et al., 2012.
+Such models can present difficult optimization problems because they are
+severely overparameterized and inevitably converge on the boundary of
+the allowable parameter space. Whether or not it is sensible to compare
+results on such extreme cases is not clear.
+
+The \texttt{SBPLX} (subplex) algorithm, which is an enhancement of
+\texttt{Nelder\_Mead}, does better in these cases but is still rather
+slow.
+
+ \begin{verbatim}
+        | Row | opt              | dsname | excess  | time    | reltime | np | n    |
+        |-----|------------------|--------|---------|---------|---------|----|------|
+        | 1   | "NLOPT\_LN\_SBPLX" | "gb12" | 0.82219 | 3.813   | 18.9228 | 20 | 512  |
+        | 2   | "NLOPT\_LN\_SBPLX" | "kb07" | 4.96546 | 564.688 | 133.308 | 72 | 1790 |
+\end{verbatim}
+
+    By comparison, the \texttt{LN\_BOBYQA} algorithm converges quite rapidly
+on the \texttt{kb07} models.
+
+\begin{Verbatim}
+        | Row | dsname | excess  | objective | time   | np |
+        |-----|--------|---------|-----------|--------|----|
+        | 1   | "kb07" | 0.01695 | 28586.3   | 4.236  | 72 |
+        | 2   | "kb07" | 0.0     | 28670.9   | 0.7033 | 16 |
+\end{Verbatim}
 

### 2015-08-04 inst/doc/Paper.tex
Filled in later sections for a very, very rough outline.  NOTE .Rnw will still not compile on Rstudio, but pdfis still created correctly

@@
 
+%%% Note that because of this table, rstudio cannot compile pdf because of margins,
+% but pdf is still available (and compiled properly) as seperate file.
 \begin{verbatim}

### 2015-08-05 inst/doc/Paper.Rnw
Removed problematic tables, reformatted some tables with xtable

--- 2015-08-03 inst/doc/Paper.Rnw
@@
 \documentclass[article]{jss}
-
+\usepackage{longtable}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 res = optdir("../JSON");
-
 @
 
-\begin{verbatim}
-  | Row | opt        | dsname          | n      | np | excess  | time    | reltime|    
-  |-----|------------|-----------------|--------|----|---------|---------|-------|
-  | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  | 27.5171|
-  | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 13.9236|
-  | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 10.8009|
-  | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  | 11.7005|
-  | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  | 13.6969|
-  | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  | 12.608|
-  | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  | 23.2699|
-  | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  | 18.8753|
-  | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  | 15.6477|
-  | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  | 22.0827|
-  | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
-  .
-  .
-  .
-  | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
-  | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  | 8.6481|
-  | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  | 3.7438|
-  | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   | 5.0422|
-  | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 | 9.2732|
-  | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  | 15.9582|
-  | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  | 1.1034|
-  | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  | 5.3039|
-  | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  | 8.6177|
-  | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  | 15.3763|
-  | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  | 1.133|
-  | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   | 2.3154|
+\begin{table}[htbp]
+\centering
+\begin{tabular}{rllrrrrrr}
+  \hline
+ & opt & dsname & n & p & np & excess & time & reltime \\ 
+  \hline
+1 & LD\_CCSAQ & Alfalfa &  72 &  12 &   1 & 0.00 & 0.00 & 1.13 \\ 
+  2 & LD\_CCSAQ & AvgDailyGain &  32 &   8 &   1 & 0.00 & 0.00 & 0.82 \\ 
+  3 & LD\_CCSAQ & AvgDailyGain &  32 &   5 &   1 & 0.00 & 0.00 & 0.93 \\ 
+  4 & LD\_CCSAQ & BIB &  24 &   8 &   1 & 0.00 & 0.00 & 0.79 \\ 
+  5 & LD\_CCSAQ & Bond &  21 &   3 &   1 & 0.00 & 0.00 & 1.03 \\ 
+  6 & LD\_CCSAQ & bs10 & 1104 &   4 &  20 & 0.00 & 1.10 & 4.44 \\ 
+  7 & LD\_CCSAQ & bs10 & 1104 &   4 &   8 & 39.99 & 0.04 & 0.56 \\ 
+  8 & LD\_CCSAQ & cake & 270 &  18 &   1 & 0.00 & 0.00 & 1.37 \\ 
+  9 & LD\_CCSAQ & Cultivation &  24 &   6 &   1 & 0.00 & 0.00 & 0.99 \\ 
+  10 & LD\_CCSAQ & Demand &  77 &   5 &   2 & 3.22 & 0.01 & 0.81 \\ 
+  11 & LD\_CCSAQ & dialectNL & 225866 &   4 &   6 & 0.00 & 6.99 & 3.99 \\ 
+  12 & LD\_CCSAQ & Dyestuff2 &  30 &   1 &   1 & 0.00 & 0.00 & 0.68 \\ 
+  13 & LD\_CCSAQ & Dyestuff &  30 &   1 &   1 & 0.00 & 0.00 & 1.02 \\ 
+  14 & LD\_CCSAQ & ergoStool &  36 &   4 &   1 & 0.00 & 0.00 & 1.10 \\ 
+  15 & LD\_CCSAQ & Exam & 4059 &   5 &   1 & 0.00 & 0.01 & 1.19 \\ 
+  16 & LD\_CCSAQ & Exam & 4059 &   6 &   1 & 0.00 & 0.01 & 1.32 \\ 
+  17 & LD\_CCSAQ & Gasoline &  32 &   2 &   1 & 0.00 & 0.00 & 0.82 \\ 
+  18 & LD\_CCSAQ & gb12 & 512 &   8 &  20 & 0.00 & 3.52 & 17.48 \\ 
+  19 & LD\_CCSAQ & gb12 & 512 &   8 &   8 & 103.18 & 0.02 & 0.56 \\ 
+  20 & LD\_CCSAQ & HR & 120 &   7 &   3 & 0.00 & 0.01 & 1.22 \\ 
+  21 & LD\_CCSAQ & Hsb82 & 7185 &   5 &   1 & 192.73 & 0.01 & 0.49 \\ 
+  22 & LD\_CCSAQ & IncBlk &  24 &   8 &   1 & 0.56 & 0.00 & 0.66 \\ 
+  23 & LD\_CCSAQ & kb07 & 1790 &   8 &  72 & 8.21 & 17.47 & 4.12 \\ 
+  24 & LD\_CCSAQ & Mississippi &  37 &   3 &   1 & 0.93 & 0.00 & 0.67 \\ 
+  25 & LD\_CCSAQ & mm0 & 69588 &   4 &   6 & 0.00 & 4.83 & 4.40 \\ 
+  26 & LD\_CCSAQ & Oxboys & 234 &   2 &   3 & 136.79 & 0.02 & 0.71 \\ 
+  27 & LD\_CCSAQ & PBIB &  60 &  15 &   1 & 0.00 & 0.00 & 1.00 \\ 
+  28 & LD\_CCSAQ & Penicillin & 144 &   1 &   2 & 0.00 & 0.01 & 4.94 \\ 
+  29 & LD\_CCSAQ & Semiconductor &  48 &  16 &   1 & 0.00 & 0.00 & 1.05 \\ 
+  30 & LD\_CCSAQ & SIMS & 3691 &   2 &   3 & 3.61 & 0.13 & 0.99 \\ 
+  31 & LD\_CCSAQ & sleepstudy & 180 &   2 &   3 & 0.00 & 0.01 & 1.75 \\ 
+  32 & LD\_CCSAQ & sleepstudy & 180 &   2 &   2 & 0.00 & 0.00 & 1.13 \\ 
+  33 & LD\_CCSAQ & TeachingII &  96 &  12 &   1 & 0.00 & 0.00 & 0.93 \\ 
+  34 & LD\_CCSAQ & Weights & 399 &   6 &   3 & 99.13 & 0.02 & 0.51 \\ 
+  35 & LD\_CCSAQ & WWheat &  60 &   2 &   3 & 18.20 & 0.01 & 0.68 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-\end{verbatim}
+The \texttt{time} column is the time in seconds to converge. The
+\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
+optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+For \proglang{Julia} the time column is the time in seconds to converge. The reltime column is the time relative to the LN_BOBYQA optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{rllrrrrrr}
+  \hline
+ & opt & dsname & n & p & np & excess & time & reltime \\ 
+  \hline
+560 & NLOPT\_LN\_BOBYQA & Alfalfa &  72 &  12 &   1 & 0.00 & 0.04 & 27.52 \\ 
+  561 & NLOPT\_LN\_BOBYQA & Animal &  20 &   1 &   2 & 0.00 & 0.02 & 13.92 \\ 
+  562 & NLOPT\_LN\_BOBYQA & Assay &  60 &  31 &   2 & 0.00 & 0.03 & 10.80 \\ 
+  563 & NLOPT\_LN\_BOBYQA & AvgDailyGain &  32 &   8 &   1 & 0.00 & 0.02 & 11.70 \\ 
+  564 & NLOPT\_LN\_BOBYQA & AvgDailyGain &  32 &   5 &   1 & 0.00 & 0.02 & 13.70 \\ 
+  565 & NLOPT\_LN\_BOBYQA & BIB &  24 &   8 &   1 & 0.00 & 0.02 & 12.61 \\ 
+  566 & NLOPT\_LN\_BOBYQA & Bond &  21 &   3 &   1 & 0.00 & 0.02 & 23.27 \\ 
+  567 & NLOPT\_LN\_BOBYQA & bs10 & 1104 &   4 &  20 & 0.00 & 4.66 & 18.88 \\ 
+  568 & NLOPT\_LN\_BOBYQA & bs10 & 1104 &   4 &   8 & 0.00 & 1.06 & 15.65 \\ 
+  569 & NLOPT\_LN\_BOBYQA & cake & 270 &  18 &   1 & 0.00 & 0.05 & 22.08 \\ 
+  570 & NLOPT\_LN\_BOBYQA & Chem97 & 31022 &   1 &   2 & 0.00 & 0.63 & 3.94 \\ 
+  571 & NLOPT\_LN\_BOBYQA & Chem97 & 31022 &   2 &   2 & 0.00 & 0.56 & 3.60 \\ 
+  572 & NLOPT\_LN\_BOBYQA & Cultivation &  24 &   6 &   1 & 0.00 & 0.02 & 23.10 \\ 
+  573 & NLOPT\_LN\_BOBYQA & d3 & 130418 &   2 &   9 & 0.00 & 231.30 & 2.16 \\ 
+  574 & NLOPT\_LN\_BOBYQA & Demand &  77 &   5 &   2 & 0.00 & 0.03 & 3.81 \\ 
+  575 & NLOPT\_LN\_BOBYQA & dialectNL & 225866 &   4 &   6 & 0.00 & 16.88 & 9.64 \\ 
+  576 & NLOPT\_LN\_BOBYQA & Dyestuff2 &  30 &   1 &   1 & 0.00 & 0.05 & 62.58 \\ 
+  577 & NLOPT\_LN\_BOBYQA & Dyestuff &  30 &   1 &   1 & 0.00 & 0.02 & 24.87 \\ 
+  578 & NLOPT\_LN\_BOBYQA & egsingle & 7230 &   5 &   2 & 0.00 & 0.17 & 2.39 \\ 
+  579 & NLOPT\_LN\_BOBYQA & ergoStool &  36 &   4 &   1 & 0.00 & 0.02 & 17.25 \\ 
+  580 & NLOPT\_LN\_BOBYQA & Exam & 4059 &   5 &   1 & 0.00 & 0.05 & 5.40 \\ 
+  581 & NLOPT\_LN\_BOBYQA & Exam & 4059 &   6 &   1 & 0.00 & 0.05 & 6.33 \\ 
+  582 & NLOPT\_LN\_BOBYQA & Gasoline &  32 &   2 &   1 & 0.00 & 0.02 & 14.30 \\ 
+  583 & NLOPT\_LN\_BOBYQA & gb12 & 512 &   8 &  20 & 0.00 & 1.87 & 9.28 \\ 
+  584 & NLOPT\_LN\_BOBYQA & gb12 & 512 &   8 &   8 & 0.00 & 0.42 & 10.77 \\ 
+  585 & NLOPT\_LN\_BOBYQA & Genetics &  60 &   5 &   2 & 0.00 & 0.03 & 9.60 \\ 
+  586 & NLOPT\_LN\_BOBYQA & HR & 120 &   7 &   3 & 0.00 & 0.02 & 3.29 \\ 
+  587 & NLOPT\_LN\_BOBYQA & Hsb82 & 7185 &   5 &   1 & 0.00 & 0.07 & 3.52 \\ 
+  588 & NLOPT\_LN\_BOBYQA & IncBlk &  24 &   8 &   1 & 0.00 & 0.02 & 12.89 \\ 
+  589 & NLOPT\_LN\_BOBYQA & InstEval & 73421 &  28 &   2 & 0.00 & 10.06 & 4.28 \\ 
+  590 & NLOPT\_LN\_BOBYQA & InstEval & 73421 &   2 &   3 & 0.00 & 15.82 & 3.63 \\ 
+  591 & NLOPT\_LN\_BOBYQA & kb07 & 1790 &   8 &  72 & 0.01 & 311.03 & 73.43 \\ 
+  592 & NLOPT\_LN\_BOBYQA & kb07 & 1790 &   8 &  16 & 0.00 & 18.50 & 26.31 \\ 
+  593 & NLOPT\_LN\_BOBYQA & Mississippi &  37 &   3 &   1 & 0.00 & 0.02 & 21.53 \\ 
+  594 & NLOPT\_LN\_BOBYQA & mm0 & 69588 &   4 &   6 & 0.00 & 26.20 & 23.88 \\ 
+  595 & NLOPT\_LN\_BOBYQA & Oxboys & 234 &   2 &   3 & 0.00 & 0.03 & 1.35 \\ 
+  596 & NLOPT\_LN\_BOBYQA & Pastes &  60 &   1 &   2 & 0.00 & 0.02 & 9.51 \\ 
+  597 & NLOPT\_LN\_BOBYQA & PBIB &  60 &  15 &   1 & 0.00 & 0.02 & 13.14 \\ 
+  598 & NLOPT\_LN\_BOBYQA & Penicillin & 144 &   1 &   2 & 0.00 & 0.02 & 8.65 \\ 
+  599 & NLOPT\_LN\_BOBYQA & Poems & 275996 &   3 &   3 & 0.00 & 21.31 & 3.74 \\ 
+  600 & NLOPT\_LN\_BOBYQA & ScotsSec & 3435 &   4 &   2 & 0.00 & 0.08 & 5.04 \\ 
+  601 & NLOPT\_LN\_BOBYQA & Semi2 &  72 &   2 &   3 & 0.00 & 0.03 & 9.27 \\ 
+  602 & NLOPT\_LN\_BOBYQA & Semiconductor &  48 &  16 &   1 & 0.00 & 0.02 & 15.96 \\ 
+  603 & NLOPT\_LN\_BOBYQA & SIMS & 3691 &   2 &   3 & 0.00 & 0.15 & 1.10 \\ 
+  604 & NLOPT\_LN\_BOBYQA & sleepstudy & 180 &   2 &   3 & 0.00 & 0.04 & 5.30 \\ 
+  605 & NLOPT\_LN\_BOBYQA & sleepstudy & 180 &   2 &   2 & 0.00 & 0.02 & 8.62 \\ 
+  606 & NLOPT\_LN\_BOBYQA & TeachingII &  96 &  12 &   1 & 0.00 & 0.02 & 15.38 \\ 
+  607 & NLOPT\_LN\_BOBYQA & Weights & 399 &   6 &   3 & 0.00 & 0.04 & 1.13 \\ 
+  608 & NLOPT\_LN\_BOBYQA & WWheat &  60 &   2 &   3 & 0.00 & 0.03 & 2.32 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
 \subsection[Convergence]{Proportion Converged}
- Etiam placerat commodo dapibus.
-Vivamus a porta dui.
-Phasellus accumsan sollicitudin cursus.
-In nec lorem sit amet purus vulputate suscipit quis ac lectus.
-Aliquam sollicitudin malesuada blandit.
-Aliquam sed nisi accumsan, pulvinar diam eget, facilisis urna.
-Fusce a dignissim ipsum.
-Integer maximus, felis nec lobortis vulputate, nunc leo blandit sem, id suscipit velit leo in purus.
-Pellentesque vel arcu sit amet nulla tincidunt ornare.
-Mauris sodales lectus eget est laoreet lobortis.
+
+The most important question regarding the optimizers is whether or not they have converged to the global optimum.
+We cannot test this directly.
+Instead we use a "crowd-sourced" criterion based on the minimum objective achieved by any of the algorithms.
+The difference between the objective achieved by a particular algorithm and this minimum is called the excess.
+In the summaries \texttt{excess} is rounded to 5 digits after the decimal so the minimum non-zero \texttt{excess} is $10^{-5}$.
+
+  If we wish to declare ``converged'' or ``not converged'' according to
+the excess objective value we must establish a threshold. An absolute
+threshold seems reasonable because the objective, negative twice the
+log-likelihood, is on a scale where differences in this objective are
+compared to a $\chi^2$ random variable. Thus an excess of $10^{-9}$ or
+even $10^{-5}$ is negligible.
+
+For each optimizer we can examine which of the data set/model
+combinations resulted in an excess greater than a threshold.
+
+    At this threshold the most reliable algorithm in Julia is
+\texttt{LN\_BOBYQA}. In \proglang{R} the most reliable algorithms are
+\texttt{NLOPT\_LN\_BOBYQA}, \texttt{optimx:L-BFGS-B} and
+\texttt{optimx:nlminb}. It is interesting that nlminb is reliable as I
+felt that it wasn't converging well when it was the default optimizer in
+\pkg{lmer}.
+
+Interestingly, the derviative-based algorithms in \texttt{NLopt} were
+not as reliable as the derivative-free algorithms. The most likely
+explanation is that I don't have the gradient coded properly.
+
+The Nelder-Mead simplex algorithm did not perform well, failing on 8 out
+of 48 cases. For many of these the value at which convergence was
+declared was far from the optimum.
+
+ The \texttt{Nelder\_Mead} algorithm, either in the native form in
+\pkg{lmer} or in the \texttt{NLopt} implementation performed poorly
+on those cases with many parameters to optimize. It was both unreliable
+and slow, taking over 45 minutes to reach a spurious optimum on the
+``maximal'' model (in the sense of Barr et al., 2012) for the
+\texttt{kb07} data from Kronmueller and Barr (2007). This is not
+terribly surprising given that the model is horribly overparameterized,
+but still it shows that this algorithm is not a good choice in these
+cases.
+
+We note in passing that all the models involving fitting 20 or more
+parameters are ``maximal'' models in the sense of Barr et al., 2012.
+Such models can present difficult optimization problems because they are
+severely overparameterized and inevitably converge on the boundary of
+the allowable parameter space. Whether or not it is sensible to compare
+results on such extreme cases is not clear.
+
+The \texttt{SBPLX} (subplex) algorithm, which is an enhancement of
+\texttt{Nelder\_Mead}, does better in these cases but is still rather
+slow.
+
+    By comparison, the \texttt{LN\_BOBYQA} algorithm converges quite rapidly
+on the \texttt{kb07} models.
 

### 2015-08-05 inst/doc/Paper.tex
Removed problematic tables, reformatted some tables with xtable

--- 2015-08-03 inst/doc/Paper.tex
@@
 > res = optdir("../JSON");
-> 
 \end{Sinput}
@@
 
-\begin{verbatim}
-  | Row | opt        | dsname          | n      | np | excess  | time    | reltime|    
-  |-----|------------|-----------------|--------|----|---------|---------|-------|
-  | 1   | "LD\_CCSAQ" | "Alfalfa"       | 72     | 1  | 0.0     | 0.0017  | 27.5171|
-  | 2   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 13.9236|
-  | 3   | "LD\_CCSAQ" | "AvgDailyGain"  | 32     | 1  | 0.0     | 0.0014  | 10.8009|
-  | 4   | "LD\_CCSAQ" | "BIB"           | 24     | 1  | 0.0     | 0.0013  | 11.7005|
-  | 5   | "LD\_CCSAQ" | "Bond"          | 21     | 1  | 0.0     | 0.0009  | 13.6969|
-  | 6   | "LD\_CCSAQ" | "bs10"          | 1104   | 20 | 0.0     | 1.0958  | 12.608|
-  | 7   | "LD\_CCSAQ" | "bs10"          | 1104   | 8  | 39.9948 | 0.0375  | 23.2699|
-  | 8   | "LD\_CCSAQ" | "cake"          | 270    | 1  | 0.0     | 0.0033  | 18.8753|
-  | 9   | "LD\_CCSAQ" | "Cultivation"   | 24     | 1  | 0.0     | 0.0009  | 15.6477|
-  | 10  | "LD\_CCSAQ" | "Demand"        | 77     | 2  | 3.21928 | 0.0055  | 22.0827|
-  | 11  | "LD\_CCSAQ" | "dialectNL"     | 225866 | 6  | 0.0     | 6.9896  | 3.942|
-  .
-  .
-  .
-  | 19  | "LD\_CCSAQ" | "gb12"          | 512    | 8  | 103.176 | 0.0218  | 13.1358|
-  | 20  | "LD\_CCSAQ" | "HR"            | 120    | 3  | 0.0     | 0.0089  | 8.6481|
-  | 21  | "LD\_CCSAQ" | "Hsb82"         | 7185   | 1  | 192.73  | 0.0102  | 3.7438|
-  | 22  | "LD\_CCSAQ" | "IncBlk"        | 24     | 1  | 0.55726 | 0.001   | 5.0422|
-  | 23  | "LD\_CCSAQ" | "kb07"          | 1790   | 72 | 8.20739 | 17.4698 | 9.2732|
-  | 24  | "LD\_CCSAQ" | "Mississippi"   | 37     | 1  | 0.93471 | 0.0006  | 15.9582|
-  | 25  | "LD\_CCSAQ" | "mm0"           | 69588  | 6  | 0.0     | 4.8286  | 1.1034|
-  | 26  | "LD\_CCSAQ" | "Oxboys"        | 234    | 3  | 136.788 | 0.0169  | 5.3039|
-  | 27  | "LD\_CCSAQ" | "PBIB"          | 60     | 1  | 0.0     | 0.0014  | 8.6177|
-  | 28  | "LD\_CCSAQ" | "Penicillin"    | 144    | 2  | 0.0     | 0.0131  | 15.3763|
-  | 29  | "LD\_CCSAQ" | "Semiconductor" | 48     | 1  | 0.0     | 0.0012  | 1.133|
-  | 30  | "LD\_CCSAQ" | "SIMS"          | 3691   | 3  | 3.60856 | 0.134   | 2.3154|
+\begin{table}[htbp]
+\centering
+\begin{tabular}{rllrrrrrr}
+  \hline
+ & opt & dsname & n & p & np & excess & time & reltime \\ 
+  \hline
+1 & LD\_CCSAQ & Alfalfa &  72 &  12 &   1 & 0.00 & 0.00 & 1.13 \\ 
+  2 & LD\_CCSAQ & AvgDailyGain &  32 &   8 &   1 & 0.00 & 0.00 & 0.82 \\ 
+  3 & LD\_CCSAQ & AvgDailyGain &  32 &   5 &   1 & 0.00 & 0.00 & 0.93 \\ 
+  4 & LD\_CCSAQ & BIB &  24 &   8 &   1 & 0.00 & 0.00 & 0.79 \\ 
+  5 & LD\_CCSAQ & Bond &  21 &   3 &   1 & 0.00 & 0.00 & 1.03 \\ 
+  6 & LD\_CCSAQ & bs10 & 1104 &   4 &  20 & 0.00 & 1.10 & 4.44 \\ 
+  7 & LD\_CCSAQ & bs10 & 1104 &   4 &   8 & 39.99 & 0.04 & 0.56 \\ 
+  8 & LD\_CCSAQ & cake & 270 &  18 &   1 & 0.00 & 0.00 & 1.37 \\ 
+  9 & LD\_CCSAQ & Cultivation &  24 &   6 &   1 & 0.00 & 0.00 & 0.99 \\ 
+  10 & LD\_CCSAQ & Demand &  77 &   5 &   2 & 3.22 & 0.01 & 0.81 \\ 
+  11 & LD\_CCSAQ & dialectNL & 225866 &   4 &   6 & 0.00 & 6.99 & 3.99 \\ 
+  12 & LD\_CCSAQ & Dyestuff2 &  30 &   1 &   1 & 0.00 & 0.00 & 0.68 \\ 
+  13 & LD\_CCSAQ & Dyestuff &  30 &   1 &   1 & 0.00 & 0.00 & 1.02 \\ 
+  14 & LD\_CCSAQ & ergoStool &  36 &   4 &   1 & 0.00 & 0.00 & 1.10 \\ 
+  15 & LD\_CCSAQ & Exam & 4059 &   5 &   1 & 0.00 & 0.01 & 1.19 \\ 
+  16 & LD\_CCSAQ & Exam & 4059 &   6 &   1 & 0.00 & 0.01 & 1.32 \\ 
+  17 & LD\_CCSAQ & Gasoline &  32 &   2 &   1 & 0.00 & 0.00 & 0.82 \\ 
+  18 & LD\_CCSAQ & gb12 & 512 &   8 &  20 & 0.00 & 3.52 & 17.48 \\ 
+  19 & LD\_CCSAQ & gb12 & 512 &   8 &   8 & 103.18 & 0.02 & 0.56 \\ 
+  20 & LD\_CCSAQ & HR & 120 &   7 &   3 & 0.00 & 0.01 & 1.22 \\ 
+  21 & LD\_CCSAQ & Hsb82 & 7185 &   5 &   1 & 192.73 & 0.01 & 0.49 \\ 
+  22 & LD\_CCSAQ & IncBlk &  24 &   8 &   1 & 0.56 & 0.00 & 0.66 \\ 
+  23 & LD\_CCSAQ & kb07 & 1790 &   8 &  72 & 8.21 & 17.47 & 4.12 \\ 
+  24 & LD\_CCSAQ & Mississippi &  37 &   3 &   1 & 0.93 & 0.00 & 0.67 \\ 
+  25 & LD\_CCSAQ & mm0 & 69588 &   4 &   6 & 0.00 & 4.83 & 4.40 \\ 
+  26 & LD\_CCSAQ & Oxboys & 234 &   2 &   3 & 136.79 & 0.02 & 0.71 \\ 
+  27 & LD\_CCSAQ & PBIB &  60 &  15 &   1 & 0.00 & 0.00 & 1.00 \\ 
+  28 & LD\_CCSAQ & Penicillin & 144 &   1 &   2 & 0.00 & 0.01 & 4.94 \\ 
+  29 & LD\_CCSAQ & Semiconductor &  48 &  16 &   1 & 0.00 & 0.00 & 1.05 \\ 
+  30 & LD\_CCSAQ & SIMS & 3691 &   2 &   3 & 3.61 & 0.13 & 0.99 \\ 
+  31 & LD\_CCSAQ & sleepstudy & 180 &   2 &   3 & 0.00 & 0.01 & 1.75 \\ 
+  32 & LD\_CCSAQ & sleepstudy & 180 &   2 &   2 & 0.00 & 0.00 & 1.13 \\ 
+  33 & LD\_CCSAQ & TeachingII &  96 &  12 &   1 & 0.00 & 0.00 & 0.93 \\ 
+  34 & LD\_CCSAQ & Weights & 399 &   6 &   3 & 99.13 & 0.02 & 0.51 \\ 
+  35 & LD\_CCSAQ & WWheat &  60 &   2 &   3 & 18.20 & 0.01 & 0.68 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-\end{verbatim}
+The \texttt{time} column is the time in seconds to converge. The
+\texttt{reltime} column is the time relative to the \texttt{LN\_BOBYQA}
+optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+For \proglang{Julia} the time column is the time in seconds to converge. The reltime column is the time relative to the LN_BOBYQA optimizer in the \pkg{MixedModels} package for \proglang{Julia}.
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{rllrrrrrr}
+  \hline
+ & opt & dsname & n & p & np & excess & time & reltime \\ 
+  \hline
+560 & NLOPT\_LN\_BOBYQA & Alfalfa &  72 &  12 &   1 & 0.00 & 0.04 & 27.52 \\ 
+  561 & NLOPT\_LN\_BOBYQA & Animal &  20 &   1 &   2 & 0.00 & 0.02 & 13.92 \\ 
+  562 & NLOPT\_LN\_BOBYQA & Assay &  60 &  31 &   2 & 0.00 & 0.03 & 10.80 \\ 
+  563 & NLOPT\_LN\_BOBYQA & AvgDailyGain &  32 &   8 &   1 & 0.00 & 0.02 & 11.70 \\ 
+  564 & NLOPT\_LN\_BOBYQA & AvgDailyGain &  32 &   5 &   1 & 0.00 & 0.02 & 13.70 \\ 
+  565 & NLOPT\_LN\_BOBYQA & BIB &  24 &   8 &   1 & 0.00 & 0.02 & 12.61 \\ 
+  566 & NLOPT\_LN\_BOBYQA & Bond &  21 &   3 &   1 & 0.00 & 0.02 & 23.27 \\ 
+  567 & NLOPT\_LN\_BOBYQA & bs10 & 1104 &   4 &  20 & 0.00 & 4.66 & 18.88 \\ 
+  568 & NLOPT\_LN\_BOBYQA & bs10 & 1104 &   4 &   8 & 0.00 & 1.06 & 15.65 \\ 
+  569 & NLOPT\_LN\_BOBYQA & cake & 270 &  18 &   1 & 0.00 & 0.05 & 22.08 \\ 
+  570 & NLOPT\_LN\_BOBYQA & Chem97 & 31022 &   1 &   2 & 0.00 & 0.63 & 3.94 \\ 
+  571 & NLOPT\_LN\_BOBYQA & Chem97 & 31022 &   2 &   2 & 0.00 & 0.56 & 3.60 \\ 
+  572 & NLOPT\_LN\_BOBYQA & Cultivation &  24 &   6 &   1 & 0.00 & 0.02 & 23.10 \\ 
+  573 & NLOPT\_LN\_BOBYQA & d3 & 130418 &   2 &   9 & 0.00 & 231.30 & 2.16 \\ 
+  574 & NLOPT\_LN\_BOBYQA & Demand &  77 &   5 &   2 & 0.00 & 0.03 & 3.81 \\ 
+  575 & NLOPT\_LN\_BOBYQA & dialectNL & 225866 &   4 &   6 & 0.00 & 16.88 & 9.64 \\ 
+  576 & NLOPT\_LN\_BOBYQA & Dyestuff2 &  30 &   1 &   1 & 0.00 & 0.05 & 62.58 \\ 
+  577 & NLOPT\_LN\_BOBYQA & Dyestuff &  30 &   1 &   1 & 0.00 & 0.02 & 24.87 \\ 
+  578 & NLOPT\_LN\_BOBYQA & egsingle & 7230 &   5 &   2 & 0.00 & 0.17 & 2.39 \\ 
+  579 & NLOPT\_LN\_BOBYQA & ergoStool &  36 &   4 &   1 & 0.00 & 0.02 & 17.25 \\ 
+  580 & NLOPT\_LN\_BOBYQA & Exam & 4059 &   5 &   1 & 0.00 & 0.05 & 5.40 \\ 
+  581 & NLOPT\_LN\_BOBYQA & Exam & 4059 &   6 &   1 & 0.00 & 0.05 & 6.33 \\ 
+  582 & NLOPT\_LN\_BOBYQA & Gasoline &  32 &   2 &   1 & 0.00 & 0.02 & 14.30 \\ 
+  583 & NLOPT\_LN\_BOBYQA & gb12 & 512 &   8 &  20 & 0.00 & 1.87 & 9.28 \\ 
+  584 & NLOPT\_LN\_BOBYQA & gb12 & 512 &   8 &   8 & 0.00 & 0.42 & 10.77 \\ 
+  585 & NLOPT\_LN\_BOBYQA & Genetics &  60 &   5 &   2 & 0.00 & 0.03 & 9.60 \\ 
+  586 & NLOPT\_LN\_BOBYQA & HR & 120 &   7 &   3 & 0.00 & 0.02 & 3.29 \\ 
+  587 & NLOPT\_LN\_BOBYQA & Hsb82 & 7185 &   5 &   1 & 0.00 & 0.07 & 3.52 \\ 
+  588 & NLOPT\_LN\_BOBYQA & IncBlk &  24 &   8 &   1 & 0.00 & 0.02 & 12.89 \\ 
+  589 & NLOPT\_LN\_BOBYQA & InstEval & 73421 &  28 &   2 & 0.00 & 10.06 & 4.28 \\ 
+  590 & NLOPT\_LN\_BOBYQA & InstEval & 73421 &   2 &   3 & 0.00 & 15.82 & 3.63 \\ 
+  591 & NLOPT\_LN\_BOBYQA & kb07 & 1790 &   8 &  72 & 0.01 & 311.03 & 73.43 \\ 
+  592 & NLOPT\_LN\_BOBYQA & kb07 & 1790 &   8 &  16 & 0.00 & 18.50 & 26.31 \\ 
+  593 & NLOPT\_LN\_BOBYQA & Mississippi &  37 &   3 &   1 & 0.00 & 0.02 & 21.53 \\ 
+  594 & NLOPT\_LN\_BOBYQA & mm0 & 69588 &   4 &   6 & 0.00 & 26.20 & 23.88 \\ 
+  595 & NLOPT\_LN\_BOBYQA & Oxboys & 234 &   2 &   3 & 0.00 & 0.03 & 1.35 \\ 
+  596 & NLOPT\_LN\_BOBYQA & Pastes &  60 &   1 &   2 & 0.00 & 0.02 & 9.51 \\ 
+  597 & NLOPT\_LN\_BOBYQA & PBIB &  60 &  15 &   1 & 0.00 & 0.02 & 13.14 \\ 
+  598 & NLOPT\_LN\_BOBYQA & Penicillin & 144 &   1 &   2 & 0.00 & 0.02 & 8.65 \\ 
+  599 & NLOPT\_LN\_BOBYQA & Poems & 275996 &   3 &   3 & 0.00 & 21.31 & 3.74 \\ 
+  600 & NLOPT\_LN\_BOBYQA & ScotsSec & 3435 &   4 &   2 & 0.00 & 0.08 & 5.04 \\ 
+  601 & NLOPT\_LN\_BOBYQA & Semi2 &  72 &   2 &   3 & 0.00 & 0.03 & 9.27 \\ 
+  602 & NLOPT\_LN\_BOBYQA & Semiconductor &  48 &  16 &   1 & 0.00 & 0.02 & 15.96 \\ 
+  603 & NLOPT\_LN\_BOBYQA & SIMS & 3691 &   2 &   3 & 0.00 & 0.15 & 1.10 \\ 
+  604 & NLOPT\_LN\_BOBYQA & sleepstudy & 180 &   2 &   3 & 0.00 & 0.04 & 5.30 \\ 
+  605 & NLOPT\_LN\_BOBYQA & sleepstudy & 180 &   2 &   2 & 0.00 & 0.02 & 8.62 \\ 
+  606 & NLOPT\_LN\_BOBYQA & TeachingII &  96 &  12 &   1 & 0.00 & 0.02 & 15.38 \\ 
+  607 & NLOPT\_LN\_BOBYQA & Weights & 399 &   6 &   3 & 0.00 & 0.04 & 1.13 \\ 
+  608 & NLOPT\_LN\_BOBYQA & WWheat &  60 &   2 &   3 & 0.00 & 0.03 & 2.32 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
 \subsection[Convergence]{Proportion Converged}
- Etiam placerat commodo dapibus.
-Vivamus a porta dui.
-Phasellus accumsan sollicitudin cursus.
-In nec lorem sit amet purus vulputate suscipit quis ac lectus.
-Aliquam sollicitudin malesuada blandit.
-Aliquam sed nisi accumsan, pulvinar diam eget, facilisis urna.
-Fusce a dignissim ipsum.
-Integer maximus, felis nec lobortis vulputate, nunc leo blandit sem, id suscipit velit leo in purus.
-Pellentesque vel arcu sit amet nulla tincidunt ornare.
-Mauris sodales lectus eget est laoreet lobortis.
+
+The most important question regarding the optimizers is whether or not they have converged to the global optimum.
+We cannot test this directly.
+Instead we use a "crowd-sourced" criterion based on the minimum objective achieved by any of the algorithms.
+The difference between the objective achieved by a particular algorithm and this minimum is called the excess.
+In the summaries \texttt{excess} is rounded to 5 digits after the decimal so the minimum non-zero \texttt{excess} is $10^{-5}$.
+
+  If we wish to declare ``converged'' or ``not converged'' according to
+the excess objective value we must establish a threshold. An absolute
+threshold seems reasonable because the objective, negative twice the
+log-likelihood, is on a scale where differences in this objective are
+compared to a $\chi^2$ random variable. Thus an excess of $10^{-9}$ or
+even $10^{-5}$ is negligible.
+
+For each optimizer we can examine which of the data set/model
+combinations resulted in an excess greater than a threshold.
+
+    At this threshold the most reliable algorithm in Julia is
+\texttt{LN\_BOBYQA}. In \proglang{R} the most reliable algorithms are
+\texttt{NLOPT\_LN\_BOBYQA}, \texttt{optimx:L-BFGS-B} and
+\texttt{optimx:nlminb}. It is interesting that nlminb is reliable as I
+felt that it wasn't converging well when it was the default optimizer in
+\pkg{lmer}.
+
+Interestingly, the derviative-based algorithms in \texttt{NLopt} were
+not as reliable as the derivative-free algorithms. The most likely
+explanation is that I don't have the gradient coded properly.
+
+The Nelder-Mead simplex algorithm did not perform well, failing on 8 out
+of 48 cases. For many of these the value at which convergence was
+declared was far from the optimum.
+
+ The \texttt{Nelder\_Mead} algorithm, either in the native form in
+\pkg{lmer} or in the \texttt{NLopt} implementation performed poorly
+on those cases with many parameters to optimize. It was both unreliable
+and slow, taking over 45 minutes to reach a spurious optimum on the
+``maximal'' model (in the sense of Barr et al., 2012) for the
+\texttt{kb07} data from Kronmueller and Barr (2007). This is not
+terribly surprising given that the model is horribly overparameterized,
+but still it shows that this algorithm is not a good choice in these
+cases.
+
+We note in passing that all the models involving fitting 20 or more
+parameters are ``maximal'' models in the sense of Barr et al., 2012.
+Such models can present difficult optimization problems because they are
+severely overparameterized and inevitably converge on the boundary of
+the allowable parameter space. Whether or not it is sensible to compare
+results on such extreme cases is not clear.
+
+The \texttt{SBPLX} (subplex) algorithm, which is an enhancement of
+\texttt{Nelder\_Mead}, does better in these cases but is still rather
+slow.
+
+    By comparison, the \texttt{LN\_BOBYQA} algorithm converges quite rapidly
+on the \texttt{kb07} models.
 

### 2015-08-27 inst/doc/Paper.Rnw
Used correct chunk options and fixed table float issue

@@
 \SweaveOpts{concordance=FALSE}
+
+<<>>=
+load("~/Desktop/Summer_Sweave/Timings/data/res.rda")
+@
+
 %% include your article here, just as usual
@@
 
-\begin{table}[htbp]
-\centering
-\begin{tabular}{rllrrrrrr}
-  \hline
- & opt & dsname & n & p & np & excess & time & reltime \\ 
-  \hline
-1 & LD\_CCSAQ & Alfalfa &  72 &  12 &   1 & 0.00 & 0.00 & 1.13 \\ 
-  2 & LD\_CCSAQ & AvgDailyGain &  32 &   8 &   1 & 0.00 & 0.00 & 0.82 \\ 
-  3 & LD\_CCSAQ & AvgDailyGain &  32 &   5 &   1 & 0.00 & 0.00 & 0.93 \\ 
-  4 & LD\_CCSAQ & BIB &  24 &   8 &   1 & 0.00 & 0.00 & 0.79 \\ 
-  5 & LD\_CCSAQ & Bond &  21 &   3 &   1 & 0.00 & 0.00 & 1.03 \\ 
-  6 & LD\_CCSAQ & bs10 & 1104 &   4 &  20 & 0.00 & 1.10 & 4.44 \\ 
-  7 & LD\_CCSAQ & bs10 & 1104 &   4 &   8 & 39.99 & 0.04 & 0.56 \\ 
-  8 & LD\_CCSAQ & cake & 270 &  18 &   1 & 0.00 & 0.00 & 1.37 \\ 
-  9 & LD\_CCSAQ & Cultivation &  24 &   6 &   1 & 0.00 & 0.00 & 0.99 \\ 
-  10 & LD\_CCSAQ & Demand &  77 &   5 &   2 & 3.22 & 0.01 & 0.81 \\ 
-  11 & LD\_CCSAQ & dialectNL & 225866 &   4 &   6 & 0.00 & 6.99 & 3.99 \\ 
-  12 & LD\_CCSAQ & Dyestuff2 &  30 &   1 &   1 & 0.00 & 0.00 & 0.68 \\ 
-  13 & LD\_CCSAQ & Dyestuff &  30 &   1 &   1 & 0.00 & 0.00 & 1.02 \\ 
-  14 & LD\_CCSAQ & ergoStool &  36 &   4 &   1 & 0.00 & 0.00 & 1.10 \\ 
-  15 & LD\_CCSAQ & Exam & 4059 &   5 &   1 & 0.00 & 0.01 & 1.19 \\ 
-  16 & LD\_CCSAQ & Exam & 4059 &   6 &   1 & 0.00 & 0.01 & 1.32 \\ 
-  17 & LD\_CCSAQ & Gasoline &  32 &   2 &   1 & 0.00 & 0.00 & 0.82 \\ 
-  18 & LD\_CCSAQ & gb12 & 512 &   8 &  20 & 0.00 & 3.52 & 17.48 \\ 
-  19 & LD\_CCSAQ & gb12 & 512 &   8 &   8 & 103.18 & 0.02 & 0.56 \\ 
-  20 & LD\_CCSAQ & HR & 120 &   7 &   3 & 0.00 & 0.01 & 1.22 \\ 
-  21 & LD\_CCSAQ & Hsb82 & 7185 &   5 &   1 & 192.73 & 0.01 & 0.49 \\ 
-  22 & LD\_CCSAQ & IncBlk &  24 &   8 &   1 & 0.56 & 0.00 & 0.66 \\ 
-  23 & LD\_CCSAQ & kb07 & 1790 &   8 &  72 & 8.21 & 17.47 & 4.12 \\ 
-  24 & LD\_CCSAQ & Mississippi &  37 &   3 &   1 & 0.93 & 0.00 & 0.67 \\ 
-  25 & LD\_CCSAQ & mm0 & 69588 &   4 &   6 & 0.00 & 4.83 & 4.40 \\ 
-  26 & LD\_CCSAQ & Oxboys & 234 &   2 &   3 & 136.79 & 0.02 & 0.71 \\ 
-  27 & LD\_CCSAQ & PBIB &  60 &  15 &   1 & 0.00 & 0.00 & 1.00 \\ 
-  28 & LD\_CCSAQ & Penicillin & 144 &   1 &   2 & 0.00 & 0.01 & 4.94 \\ 
-  29 & LD\_CCSAQ & Semiconductor &  48 &  16 &   1 & 0.00 & 0.00 & 1.05 \\ 
-  30 & LD\_CCSAQ & SIMS & 3691 &   2 &   3 & 3.61 & 0.13 & 0.99 \\ 
-  31 & LD\_CCSAQ & sleepstudy & 180 &   2 &   3 & 0.00 & 0.01 & 1.75 \\ 
-  32 & LD\_CCSAQ & sleepstudy & 180 &   2 &   2 & 0.00 & 0.00 & 1.13 \\ 
-  33 & LD\_CCSAQ & TeachingII &  96 &  12 &   1 & 0.00 & 0.00 & 0.93 \\ 
-  34 & LD\_CCSAQ & Weights & 399 &   6 &   3 & 99.13 & 0.02 & 0.51 \\ 
-  35 & LD\_CCSAQ & WWheat &  60 &   2 &   3 & 18.20 & 0.01 & 0.68 \\ 
-   \hline
-\end{tabular}
-\end{table}
+<<echo=FALSE>>=
+library(xtable)
+@
+
+\begin
+{center}
+<<results=tex,echo=FALSE>>=
+print(xtable(subset(res, opt=="LD_CCSAQ")[1:8]), table.placement="!htp")
+@
+\end {center}
 

### 2015-08-27 inst/doc/Paper.Rnw
Reconfigured chunk float options, added more tables, general edits

@@
 
-\begin{table}[htbp]
-\centering
-\begin{tabular}{rllrrrrrr}
-  \hline
- & opt & dsname & n & p & np & excess & time & reltime \\ 
-  \hline
-560 & NLOPT\_LN\_BOBYQA & Alfalfa &  72 &  12 &   1 & 0.00 & 0.04 & 27.52 \\ 
-  561 & NLOPT\_LN\_BOBYQA & Animal &  20 &   1 &   2 & 0.00 & 0.02 & 13.92 \\ 
-  562 & NLOPT\_LN\_BOBYQA & Assay &  60 &  31 &   2 & 0.00 & 0.03 & 10.80 \\ 
-  563 & NLOPT\_LN\_BOBYQA & AvgDailyGain &  32 &   8 &   1 & 0.00 & 0.02 & 11.70 \\ 
-  564 & NLOPT\_LN\_BOBYQA & AvgDailyGain &  32 &   5 &   1 & 0.00 & 0.02 & 13.70 \\ 
-  565 & NLOPT\_LN\_BOBYQA & BIB &  24 &   8 &   1 & 0.00 & 0.02 & 12.61 \\ 
-  566 & NLOPT\_LN\_BOBYQA & Bond &  21 &   3 &   1 & 0.00 & 0.02 & 23.27 \\ 
-  567 & NLOPT\_LN\_BOBYQA & bs10 & 1104 &   4 &  20 & 0.00 & 4.66 & 18.88 \\ 
-  568 & NLOPT\_LN\_BOBYQA & bs10 & 1104 &   4 &   8 & 0.00 & 1.06 & 15.65 \\ 
-  569 & NLOPT\_LN\_BOBYQA & cake & 270 &  18 &   1 & 0.00 & 0.05 & 22.08 \\ 
-  570 & NLOPT\_LN\_BOBYQA & Chem97 & 31022 &   1 &   2 & 0.00 & 0.63 & 3.94 \\ 
-  571 & NLOPT\_LN\_BOBYQA & Chem97 & 31022 &   2 &   2 & 0.00 & 0.56 & 3.60 \\ 
-  572 & NLOPT\_LN\_BOBYQA & Cultivation &  24 &   6 &   1 & 0.00 & 0.02 & 23.10 \\ 
-  573 & NLOPT\_LN\_BOBYQA & d3 & 130418 &   2 &   9 & 0.00 & 231.30 & 2.16 \\ 
-  574 & NLOPT\_LN\_BOBYQA & Demand &  77 &   5 &   2 & 0.00 & 0.03 & 3.81 \\ 
-  575 & NLOPT\_LN\_BOBYQA & dialectNL & 225866 &   4 &   6 & 0.00 & 16.88 & 9.64 \\ 
-  576 & NLOPT\_LN\_BOBYQA & Dyestuff2 &  30 &   1 &   1 & 0.00 & 0.05 & 62.58 \\ 
-  577 & NLOPT\_LN\_BOBYQA & Dyestuff &  30 &   1 &   1 & 0.00 & 0.02 & 24.87 \\ 
-  578 & NLOPT\_LN\_BOBYQA & egsingle & 7230 &   5 &   2 & 0.00 & 0.17 & 2.39 \\ 
-  579 & NLOPT\_LN\_BOBYQA & ergoStool &  36 &   4 &   1 & 0.00 & 0.02 & 17.25 \\ 
-  580 & NLOPT\_LN\_BOBYQA & Exam & 4059 &   5 &   1 & 0.00 & 0.05 & 5.40 \\ 
-  581 & NLOPT\_LN\_BOBYQA & Exam & 4059 &   6 &   1 & 0.00 & 0.05 & 6.33 \\ 
-  582 & NLOPT\_LN\_BOBYQA & Gasoline &  32 &   2 &   1 & 0.00 & 0.02 & 14.30 \\ 
-  583 & NLOPT\_LN\_BOBYQA & gb12 & 512 &   8 &  20 & 0.00 & 1.87 & 9.28 \\ 
-  584 & NLOPT\_LN\_BOBYQA & gb12 & 512 &   8 &   8 & 0.00 & 0.42 & 10.77 \\ 
-  585 & NLOPT\_LN\_BOBYQA & Genetics &  60 &   5 &   2 & 0.00 & 0.03 & 9.60 \\ 
-  586 & NLOPT\_LN\_BOBYQA & HR & 120 &   7 &   3 & 0.00 & 0.02 & 3.29 \\ 
-  587 & NLOPT\_LN\_BOBYQA & Hsb82 & 7185 &   5 &   1 & 0.00 & 0.07 & 3.52 \\ 
-  588 & NLOPT\_LN\_BOBYQA & IncBlk &  24 &   8 &   1 & 0.00 & 0.02 & 12.89 \\ 
-  589 & NLOPT\_LN\_BOBYQA & InstEval & 73421 &  28 &   2 & 0.00 & 10.06 & 4.28 \\ 
-  590 & NLOPT\_LN\_BOBYQA & InstEval & 73421 &   2 &   3 & 0.00 & 15.82 & 3.63 \\ 
-  591 & NLOPT\_LN\_BOBYQA & kb07 & 1790 &   8 &  72 & 0.01 & 311.03 & 73.43 \\ 
-  592 & NLOPT\_LN\_BOBYQA & kb07 & 1790 &   8 &  16 & 0.00 & 18.50 & 26.31 \\ 
-  593 & NLOPT\_LN\_BOBYQA & Mississippi &  37 &   3 &   1 & 0.00 & 0.02 & 21.53 \\ 
-  594 & NLOPT\_LN\_BOBYQA & mm0 & 69588 &   4 &   6 & 0.00 & 26.20 & 23.88 \\ 
-  595 & NLOPT\_LN\_BOBYQA & Oxboys & 234 &   2 &   3 & 0.00 & 0.03 & 1.35 \\ 
-  596 & NLOPT\_LN\_BOBYQA & Pastes &  60 &   1 &   2 & 0.00 & 0.02 & 9.51 \\ 
-  597 & NLOPT\_LN\_BOBYQA & PBIB &  60 &  15 &   1 & 0.00 & 0.02 & 13.14 \\ 
-  598 & NLOPT\_LN\_BOBYQA & Penicillin & 144 &   1 &   2 & 0.00 & 0.02 & 8.65 \\ 
-  599 & NLOPT\_LN\_BOBYQA & Poems & 275996 &   3 &   3 & 0.00 & 21.31 & 3.74 \\ 
-  600 & NLOPT\_LN\_BOBYQA & ScotsSec & 3435 &   4 &   2 & 0.00 & 0.08 & 5.04 \\ 
-  601 & NLOPT\_LN\_BOBYQA & Semi2 &  72 &   2 &   3 & 0.00 & 0.03 & 9.27 \\ 
-  602 & NLOPT\_LN\_BOBYQA & Semiconductor &  48 &  16 &   1 & 0.00 & 0.02 & 15.96 \\ 
-  603 & NLOPT\_LN\_BOBYQA & SIMS & 3691 &   2 &   3 & 0.00 & 0.15 & 1.10 \\ 
-  604 & NLOPT\_LN\_BOBYQA & sleepstudy & 180 &   2 &   3 & 0.00 & 0.04 & 5.30 \\ 
-  605 & NLOPT\_LN\_BOBYQA & sleepstudy & 180 &   2 &   2 & 0.00 & 0.02 & 8.62 \\ 
-  606 & NLOPT\_LN\_BOBYQA & TeachingII &  96 &  12 &   1 & 0.00 & 0.02 & 15.38 \\ 
-  607 & NLOPT\_LN\_BOBYQA & Weights & 399 &   6 &   3 & 0.00 & 0.04 & 1.13 \\ 
-  608 & NLOPT\_LN\_BOBYQA & WWheat &  60 &   2 &   3 & 0.00 & 0.03 & 2.32 \\ 
-   \hline
-\end{tabular}
-\end{table}
+\begin
+{center}
+<<results=tex,echo=FALSE>>=
+print(xtable(subset(res, opt=="NLOPT_LN_BOBYQA")[1:8]), floating=FALSE, table.placement="H")
+@
+\end {center}
 
 \subsection[Convergence]{Proportion Converged}
-
 The most important question regarding the optimizers is whether or not they have converged to the global optimum.
@@
 In the summaries \texttt{excess} is rounded to 5 digits after the decimal so the minimum non-zero \texttt{excess} is $10^{-5}$.
+
+\begin
+{center}
+<<results=tex,echo=FALSE>>=
+print(xtable(subset(res, opt=="LN_BOBYQA")[c(1,2,6)]),floating=FALSE, table.placement="H")
+@
+\end {center}
 
@@
 declared was far from the optimum.
+
+<<results=tex>>=
+print(xtable(subset(res, opt=="Nelder_Mead" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+@
+
+<<results=tex>>=
+print(xtable(subset(res, opt=="NLOPT_LN_Nelder_Mead" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+@
 

### 2015-08-27 inst/doc/Paper.tex
Used correct chunk options and fixed table float issue
Reconfigured chunk float options, added more tables, general edits

@@
 \begin{document}
+
+\begin{Schunk}
+\begin{Sinput}
+> load("~/Desktop/Summer_Sweave/Timings/data/res.rda")
+\end{Sinput}
+\end{Schunk}
 
@@
 
-\begin{table}[htbp]
+\begin
+{center}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Thu Aug 27 12:42:05 2015
+\begin{table}[!htp]
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table}\end {center}
 
@@
 
-\begin{table}[htbp]
-\centering
+\begin
+{center}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Thu Aug 27 12:42:05 2015
 \begin{tabular}{rllrrrrrr}
@@
    \hline
-\end{tabular}
-\end{table}
+\end{tabular}\end {center}
 
 \subsection[Convergence]{Proportion Converged}
-
 The most important question regarding the optimizers is whether or not they have converged to the global optimum.
@@
 In the summaries \texttt{excess} is rounded to 5 digits after the decimal so the minimum non-zero \texttt{excess} is $10^{-5}$.
+
+\begin
+{center}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Thu Aug 27 12:42:05 2015
+\begin{tabular}{rllr}
+  \hline
+ & opt & dsname & excess \\ 
+  \hline
+347 & LN\_BOBYQA & Alfalfa & 0.00 \\ 
+  348 & LN\_BOBYQA & Animal & 0.00 \\ 
+  349 & LN\_BOBYQA & Assay & 0.00 \\ 
+  350 & LN\_BOBYQA & AvgDailyGain & 0.00 \\ 
+  351 & LN\_BOBYQA & AvgDailyGain & 0.00 \\ 
+  352 & LN\_BOBYQA & BIB & 0.00 \\ 
+  353 & LN\_BOBYQA & Bond & 0.00 \\ 
+  354 & LN\_BOBYQA & bs10 & 0.00 \\ 
+  355 & LN\_BOBYQA & bs10 & 0.00 \\ 
+  356 & LN\_BOBYQA & cake & 0.00 \\ 
+  357 & LN\_BOBYQA & Chem97 & 0.00 \\ 
+  358 & LN\_BOBYQA & Chem97 & 0.00 \\ 
+  359 & LN\_BOBYQA & Cultivation & 0.00 \\ 
+  360 & LN\_BOBYQA & d3 & 0.00 \\ 
+  361 & LN\_BOBYQA & Demand & 0.00 \\ 
+  362 & LN\_BOBYQA & dialectNL & 0.00 \\ 
+  363 & LN\_BOBYQA & Dyestuff2 & 0.00 \\ 
+  364 & LN\_BOBYQA & Dyestuff & 0.00 \\ 
+  365 & LN\_BOBYQA & egsingle & 0.00 \\ 
+  366 & LN\_BOBYQA & ergoStool & 0.00 \\ 
+  367 & LN\_BOBYQA & Exam & 0.00 \\ 
+  368 & LN\_BOBYQA & Exam & 0.00 \\ 
+  369 & LN\_BOBYQA & Gasoline & 0.00 \\ 
+  370 & LN\_BOBYQA & gb12 & 0.00 \\ 
+  371 & LN\_BOBYQA & gb12 & 0.00 \\ 
+  372 & LN\_BOBYQA & Genetics & 0.00 \\ 
+  373 & LN\_BOBYQA & HR & 0.00 \\ 
+  374 & LN\_BOBYQA & Hsb82 & 0.00 \\ 
+  375 & LN\_BOBYQA & IncBlk & 0.00 \\ 
+  376 & LN\_BOBYQA & InstEval & 0.00 \\ 
+  377 & LN\_BOBYQA & InstEval & 0.00 \\ 
+  378 & LN\_BOBYQA & kb07 & 0.02 \\ 
+  379 & LN\_BOBYQA & kb07 & 0.00 \\ 
+  380 & LN\_BOBYQA & Mississippi & 0.00 \\ 
+  381 & LN\_BOBYQA & mm0 & 0.00 \\ 
+  382 & LN\_BOBYQA & Oxboys & 0.00 \\ 
+  383 & LN\_BOBYQA & Pastes & 0.00 \\ 
+  384 & LN\_BOBYQA & PBIB & 0.00 \\ 
+  385 & LN\_BOBYQA & Penicillin & 0.00 \\ 
+  386 & LN\_BOBYQA & Poems & 0.00 \\ 
+  387 & LN\_BOBYQA & ScotsSec & 0.00 \\ 
+  388 & LN\_BOBYQA & Semi2 & 0.00 \\ 
+  389 & LN\_BOBYQA & Semiconductor & 0.00 \\ 
+  390 & LN\_BOBYQA & SIMS & 0.00 \\ 
+  391 & LN\_BOBYQA & sleepstudy & 0.00 \\ 
+  392 & LN\_BOBYQA & sleepstudy & 0.00 \\ 
+  393 & LN\_BOBYQA & TeachingII & 0.00 \\ 
+  394 & LN\_BOBYQA & Weights & 0.00 \\ 
+  395 & LN\_BOBYQA & WWheat & 0.00 \\ 
+   \hline
+\end{tabular}\end {center}
 
@@
 declared was far from the optimum.
+
+\begin{Schunk}
+\begin{Sinput}
+> print(xtable(subset(res, opt=="Nelder_Mead" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+\end{Sinput}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Thu Aug 27 12:42:05 2015
+\begin{tabular}{rllrrrrr}
+  \hline
+ & opt & dsname & n & np & excess & time & reltime \\ 
+  \hline
+777 & Nelder\_Mead & bs10 & 1104 &  20 & 71.39 & 145.37 & 588.68 \\ 
+  783 & Nelder\_Mead & d3 & 130418 &   9 & 317.59 & 454.52 & 4.25 \\ 
+  785 & Nelder\_Mead & dialectNL & 225866 &   6 & 181.63 & 54.54 & 31.16 \\ 
+  793 & Nelder\_Mead & gb12 & 512 &  20 & 78.71 & 38.21 & 189.60 \\ 
+  801 & Nelder\_Mead & kb07 & 1790 &  72 & 398.73 & 2825.46 & 667.02 \\ 
+  802 & Nelder\_Mead & kb07 & 1790 &  16 & 403.48 & 269.44 & 383.09 \\ 
+  803 & Nelder\_Mead & Mississippi &  37 &   1 & 0.04 & 0.02 & 20.40 \\ 
+  804 & Nelder\_Mead & mm0 & 69588 &   6 & 181.63 & 76.87 & 70.07 \\ 
+   \hline
+\end{tabular}\end{Schunk}
 

### 2015-09-01 inst/julia/retime.jl
in retime second argument can now be a directory

@@ -39,7 +39,7 @@ function retime(fnm,ofile)
             println()
         end
     end
-    open(ofile,"w") do io
+    open(isdir(ofile) ? joinpath(ofile,basename(fnm)) : ofile,"w") do io
         write(io,json(js,2))
     end
 end

### 2015-09-01 inst/doc/Paper.Rnw

@@
 <<results=tex>>=
-print(xtable(subset(res, opt=="NLOPT_LN_Nelder_Mead" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+print(xtable(subset(res, opt=="NLOPT_LN_NELDERMEAD" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
 @
@@
 
+<<results=tex>>=
+print(xtable(subset(res, opt=="NLOPT_LN_SBPLX" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+@
+
     By comparison, the \texttt{LN\_BOBYQA} algorithm converges quite rapidly
 on the \texttt{kb07} models.
+
+<<results=tex>>=
+print(xtable(subset(res, opt=="LN_BOBYQA" & dsname == "kb07")[c(1,2,5,6,7,9)]),floating=FALSE, table.placement="H")
+@
+
+\subsection[Relative Speed]{Relative Speed}
+
+We plot the time to convergence, relative to LN_BOBYQA and on a logarithmic scale, for each algorithm.
+
+PLOT
+
+Many of the cases where LN_BOBYQA is slower than other algorithms are simple problems that converge in less than 1/5 of a second for most algorithms.
+
+We will declare a data set to be non-simple if at least one of the models fit to the data took more than 0.2 seconds to convergence with LN_BOBYQA.
+
+<<results=tex>>=
+print(xtable(subset(res, opt=="LN_BOBYQA" & time >= 0.2)[c("dsname","time","n","np","models")]),floating=FALSE, table.placement="H")
+@
 

### 2015-09-09 inst/doc/Paper.Rnw
Added citations, final plot and general edits

@@
 
-<<>>=
+<<echo=FALSE>>=
 load("~/Desktop/Summer_Sweave/Timings/data/res.rda")
+library(ggplot2)
 @
@@
 @
-
-INCLUDE SUMMARY TABLE USING res.rda
 
@@
 
-PLOT
+<<fig=TRUE>>=
+res$cvg<-rep("yes",nrow(res))
+res$cvg[which(res$excess>.02)] <- "no"
+res$cvg<-as.factor(res$cvg)
+k<-ggplot(res, aes(x=log(reltime), y=opt))
+k + geom_point(aes(color=cvg))
+@
 
@@
 
-Note all optimization packages used, lmm, lme4, jsonlite,
+<<echo=FALSE,eval=FALSE>>=
+toBibtex(citation(package="lme4"))
+toBibtex(citation(package="jsonlite"))
+toBibtex(citation(package="optimx"))
+toBibtex(citation(package="minqa"))
+toBibtex(citation(package="nloptr"))
+@
+
+Douglas Bates, Katharine M. Mullen, John C. Nash and Ravi Varadhan (2014). minqa:
+  Derivative-free optimization algorithms by quadratic approximation. R package version 1.2.4.
+  http://CRAN.R-project.org/package=minqa
+
+Bates D, Maechler M, Bolker B and Walker S (2014). _lme4: Linear mixed-effects models using Eigen
+  and S4_. R package version 1.1-7, <URL: http://CRAN.R-project.org/package=lme4>.
+
+John C. Nash, Ravi Varadhan (2011). Unifying Optimization Algorithms to Aid Software System
+  Users: optimx for R. Journal of Statistical Software, 43(9), 1-14. URL
+  http://www.jstatsoft.org/v43/i09/.
+
+Jeroen Ooms (2014). The jsonlite Package: A Practical and Consistent Mapping Between JSON Data
+  and R Objects. arXiv:1403.2805 [stat.CO] URL http://arxiv.org/abs/1403.2805.
+
+Steven G. Johnson, The NLopt nonlinear-optimization package, http://ab-initio.mit.edu/nlopt
+
+(For the nloptr package they want us to go through and cite each algorithm we used)
 

### 2015-09-09 inst/doc/Paper.tex
Added citations, final plot and general edits

@@
 \begin{document}
-
-\begin{Schunk}
-\begin{Sinput}
-> load("~/Desktop/Summer_Sweave/Timings/data/res.rda")
-\end{Sinput}
-\end{Schunk}
 
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Thu Aug 27 12:42:05 2015
+% Wed Sep  9 17:45:30 2015
 \begin{table}[!htp]
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Thu Aug 27 12:42:05 2015
+% Wed Sep  9 17:45:31 2015
 \begin{tabular}{rllrrrrrr}
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Thu Aug 27 12:42:05 2015
+% Wed Sep  9 17:45:31 2015
 \begin{tabular}{rllr}
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Thu Aug 27 12:42:05 2015
+% Wed Sep  9 17:45:31 2015
 \begin{tabular}{rllrrrrr}
@@
   804 & Nelder\_Mead & mm0 & 69588 &   6 & 181.63 & 76.87 & 70.07 \\ 
+   \hline
+\end{tabular}\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+> print(xtable(subset(res, opt=="NLOPT_LN_NELDERMEAD" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+\end{Sinput}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Wed Sep  9 17:45:31 2015
+\begin{tabular}{rllrrrrr}
+  \hline
+ & opt & dsname & n & np & excess & time & reltime \\ 
+  \hline
+659 & NLOPT\_LN\_NELDERMEAD & Assay &  60 &   2 & 0.06 & 0.04 & 14.18 \\ 
+  664 & NLOPT\_LN\_NELDERMEAD & bs10 & 1104 &  20 & 0.97 & 88.29 & 357.52 \\ 
+  670 & NLOPT\_LN\_NELDERMEAD & d3 & 130418 &   9 & 100.19 & 627.93 & 5.87 \\ 
+  680 & NLOPT\_LN\_NELDERMEAD & gb12 & 512 &  20 & 0.84 & 52.41 & 260.10 \\ 
+  688 & NLOPT\_LN\_NELDERMEAD & kb07 & 1790 &  72 & 88.77 & 2839.07 & 670.23 \\ 
+  689 & NLOPT\_LN\_NELDERMEAD & kb07 & 1790 &  16 & 3.50 & 198.84 & 282.72 \\ 
    \hline
@@
 
+\begin{Schunk}
+\begin{Sinput}
+> print(xtable(subset(res, opt=="NLOPT_LN_SBPLX" & excess > 0.005)[c(1,2,3,5,6,7,8)]),floating=FALSE, table.placement="H")
+\end{Sinput}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Wed Sep  9 17:45:31 2015
+\begin{tabular}{rllrrrrr}
+  \hline
+ & opt & dsname & n & np & excess & time & reltime \\ 
+  \hline
+745 & NLOPT\_LN\_SBPLX & gb12 & 512 &  20 & 0.82 & 3.81 & 18.92 \\ 
+  753 & NLOPT\_LN\_SBPLX & kb07 & 1790 &  72 & 4.97 & 564.69 & 133.31 \\ 
+   \hline
+\end{tabular}\end{Schunk}
+
     By comparison, the \texttt{LN\_BOBYQA} algorithm converges quite rapidly
 on the \texttt{kb07} models.
+
+\begin{Schunk}
+\begin{Sinput}
+> print(xtable(subset(res, opt=="LN_BOBYQA" & dsname == "kb07")[c(1,2,5,6,7,9)]),floating=FALSE, table.placement="H")
+\end{Sinput}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Wed Sep  9 17:45:31 2015
+\begin{tabular}{rllrrrr}
+  \hline
+ & opt & dsname & np & excess & time & objective \\ 
+  \hline
+378 & LN\_BOBYQA & kb07 &  72 & 0.02 & 4.24 & 28586.33 \\ 
+  379 & LN\_BOBYQA & kb07 &  16 & 0.00 & 0.70 & 28670.91 \\ 
+   \hline
+\end{tabular}\end{Schunk}
+
+\subsection[Relative Speed]{Relative Speed}
+
+We plot the time to convergence, relative to LN_BOBYQA and on a logarithmic scale, for each algorithm.
+
+\begin{Schunk}
+\begin{Sinput}
+> res$cvg<-rep("yes",nrow(res))
+> res$cvg[which(res$excess>.02)] <- "no"
+> res$cvg<-as.factor(res$cvg)
+> k<-ggplot(res, aes(x=log(reltime), y=opt))
+> k + geom_point(aes(color=cvg))
+\end{Sinput}
+\end{Schunk}
+\includegraphics{Paper-013}
+
+Many of the cases where LN_BOBYQA is slower than other algorithms are simple problems that converge in less than 1/5 of a second for most algorithms.
+
+We will declare a data set to be non-simple if at least one of the models fit to the data took more than 0.2 seconds to convergence with LN_BOBYQA.
+
+\begin{Schunk}
+\begin{Sinput}
+> print(xtable(subset(res, opt=="LN_BOBYQA" & time >= 0.2)[c("dsname","time","n","np","models")]),floating=FALSE, table.placement="H")
+\end{Sinput}
+% latex table generated in R 3.1.3 by xtable 1.7-4 package
+% Wed Sep  9 17:45:31 2015
+\begin{tabular}{rlrrrl}
+  \hline
+ & dsname & time & n & np & models \\ 
+  \hline
+354 & bs10 & 0.25 & 1104 &  20 & Y \~{} 1 + U + V + W + ((1 + U + V + W) $|$ G) + ((1 + U + V + W) $|$ H) \\ 
+  360 & d3 & 106.94 & 130418 &   9 & Y \~{} 1 + U + ((1 + U) $|$ G) + ((1 + U) $|$ H) + ((1 + U) $|$ I) \\ 
+  362 & dialectNL & 1.75 & 225866 &   6 & Y \~{} 1 + A + U + V + W + X + Z + T + (1 $|$ G) + ((0 + V + W + X) $|$ G) + ((1 + A + Z) $|$ H) + (1 $|$ I) \\ 
+  370 & gb12 & 0.20 & 512 &  20 & Y \~{} 1 + S + T + U + V + W + X + Z + ((1 + S + U + W) $|$ G) + ((1 + S + T + V) $|$ H) \\ 
+  376 & InstEval & 2.35 & 73421 &   2 & Y \~{} 1 + I * A + (1 $|$ G) + (1 $|$ H) \\ 
+  377 & InstEval & 4.36 & 73421 &   3 & Y \~{} 1 + A + (1 $|$ G) + (1 $|$ H) + (1 $|$ I) \\ 
+  378 & kb07 & 4.24 & 1790 &  72 & Y \~{} 1 + S + T + U + V + W + X + Z + ((1 + S + T + U + V + W + X + Z) $|$ G) + ((1 + S + T + U + V + W + X + Z) $|$ H) \\ 
+  379 & kb07 & 0.70 & 1790 &  16 & Y \~{} 1 + S + T + U + V + W + X + Z + (1 $|$ G) + ((0 + S) $|$ G) + ((0 + T) $|$ G) + ((0 + U) $|$ G) + ((0 + V) $|$ G) + ((0 + W) $|$ G) + ((0 + X) $|$ G) + ((0 + Z) $|$ G) + (1 $|$ H) + ((0 + S) $|$ H) + ((0 + T) $|$ H) + ((0 + U) $|$ H) + ((0 + V) $|$ H) + ((0 + W) $|$ H) + ((0 + X) $|$ H) + ((0 + Z) $|$ H) \\
+  381 & mm0 & 1.10 & 69588 &   6 & Y \~{} 1 + A * U + ((1 + U) $|$ G) + ((1 + U) $|$ H) \\ 
+  386 & Poems & 5.69 & 275996 &   3 & Y \~{} 1 + U + V + (1 $|$ G) + (1 $|$ H) + (1 $|$ I) \\ 
+   \hline
+\end{tabular}\end{Schunk}
 
@@
 
-Note all optimization packages used, lmm, lme4, jsonlite,
+\begin{Schunk}
+\begin{Soutput}
+@Manual{,
+  title = {{lme4}: Linear mixed-effects models using Eigen and S4},
+  author = {Douglas Bates and Martin Maechler and Ben Bolker and Steven Walker},
+  year = {2014},
+  note = {R package version 1.1-7},
+  url = {http://CRAN.R-project.org/package=lme4},
+}
+
+@Misc{,
+  title = {{lme4}: Linear mixed-effects models using Eigen and S4},
+  year = {2014},
+  author = {Douglas Bates and Martin Maechler and Benjamin M. Bolker and Steven Walker},
+  note = {ArXiv e-print; submitted to \emph{Journal of Statistical Software}},
+  url = {http://arxiv.org/abs/1406.5823},
+}
+\end{Soutput}
+\begin{Soutput}
+@Article{,
+  title = {The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects},
+  author = {Jeroen Ooms},
+  journal = {arXiv:1403.2805 [stat.CO]},
+  year = {2014},
+  url = {http://arxiv.org/abs/1403.2805},
+}
+\end{Soutput}
+\begin{Soutput}
+@Article{,
+  title = {Unifying Optimization Algorithms to Aid Software System Users: {optimx} for {R}},
+  author = {John C. Nash and Ravi Varadhan},
+  journal = {Journal of Statistical Software},
+  year = {2011},
+  volume = {43},
+  number = {9},
+  pages = {1--14},
+  url = {http://www.jstatsoft.org/v43/i09/},
+}
+\end{Soutput}
+\begin{Soutput}
+@Manual{,
+  title = {minqa: Derivative-free optimization algorithms by quadratic
+approximation},
+  author = {Douglas Bates and Katharine M. Mullen and John C. Nash and Ravi Varadhan},
+  year = {2014},
+  note = {R package version 1.2.4},
+  url = {http://CRAN.R-project.org/package=minqa},
+}
+\end{Soutput}
+\begin{Soutput}
+@Article{,
+  title = {The NLopt nonlinear-optimization package},
+  author = {Steven G. Johnson},
+  year = {?},
+  journal = {?},
+  volume = {?},
+  number = {?},
+  pages = {?},
+}
+\end{Soutput}
+\end{Schunk}
 

### 2015-09-24 inst/doc/Paper.Rnw
Changes to the plot, minor edits

@@
 
+\setkeys{Gin}{width=.99\textwidth}
 <<fig=TRUE>>=
@@
 k<-ggplot(res, aes(x=log(reltime), y=opt))
-k + geom_point(aes(color=cvg))
+k + geom_point(aes(color=cvg))+scale_x_log10()
 @
@@
 
-Bates D, Maechler M, Bolker B and Walker S (2014). _lme4: Linear mixed-effects models using Eigen
+Bates D, Maechler M, Bolker B and Walker S (2014). \_lme4: Linear mixed-effects models using Eigen
   and S4_. R package version 1.1-7, <URL: http://CRAN.R-project.org/package=lme4>.

### 2015-09-24 inst/doc/Paper.tex
Changes to the plot, minor edits

@@
 
-INCLUDE SUMMARY TABLE USING res.rda
-
   The timing was repeated so that compilation time is not included in the
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Wed Sep  9 17:45:30 2015
+% Thu Sep 24 15:26:30 2015
 \begin{table}[!htp]
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Wed Sep  9 17:45:31 2015
+% Thu Sep 24 15:26:30 2015
 \begin{tabular}{rllrrrrrr}
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Wed Sep  9 17:45:31 2015
+% Thu Sep 24 15:26:30 2015
 \begin{tabular}{rllr}
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Wed Sep  9 17:45:31 2015
+% Thu Sep 24 15:26:30 2015
 \begin{tabular}{rllrrrrr}
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Wed Sep  9 17:45:31 2015
+% Thu Sep 24 15:26:30 2015
 \begin{tabular}{rllrrrr}
@@
 
+\setkeys{Gin}{width=.99\textwidth}
 \begin{Schunk}
@@
 > k<-ggplot(res, aes(x=log(reltime), y=opt))
-> k + geom_point(aes(color=cvg))
+> k + geom_point(aes(color=cvg))+scale_x_log10()
 \end{Sinput}
@@
 % latex table generated in R 3.1.3 by xtable 1.7-4 package
-% Wed Sep  9 17:45:31 2015
+% Thu Sep 24 15:26:31 2015
 \begin{tabular}{rlrrrl}
@@
 
-\begin{Schunk}
-\begin{Soutput}
-@Manual{,
-  title = {{lme4}: Linear mixed-effects models using Eigen and S4},
-  author = {Douglas Bates and Martin Maechler and Ben Bolker and Steven Walker},
-  year = {2014},
-  note = {R package version 1.1-7},
-  url = {http://CRAN.R-project.org/package=lme4},
-}
+Douglas Bates, Katharine M. Mullen, John C. Nash and Ravi Varadhan (2014). minqa:
+  Derivative-free optimization algorithms by quadratic approximation. R package version 1.2.4.
+  http://CRAN.R-project.org/package=minqa
 
-@Misc{,
-  title = {{lme4}: Linear mixed-effects models using Eigen and S4},
-  year = {2014},
-  author = {Douglas Bates and Martin Maechler and Benjamin M. Bolker and Steven Walker},
-  note = {ArXiv e-print; submitted to \emph{Journal of Statistical Software}},
-  url = {http://arxiv.org/abs/1406.5823},
-}
-\end{Soutput}
-\begin{Soutput}
-@Article{,
-  title = {The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects},
-  author = {Jeroen Ooms},
-  journal = {arXiv:1403.2805 [stat.CO]},
-  year = {2014},
-  url = {http://arxiv.org/abs/1403.2805},
-}
-\end{Soutput}
-\begin{Soutput}
-@Article{,
-  title = {Unifying Optimization Algorithms to Aid Software System Users: {optimx} for {R}},
-  author = {John C. Nash and Ravi Varadhan},
-  journal = {Journal of Statistical Software},
-  year = {2011},
-  volume = {43},
-  number = {9},
-  pages = {1--14},
-  url = {http://www.jstatsoft.org/v43/i09/},
-}
-\end{Soutput}
-\begin{Soutput}
-@Manual{,
-  title = {minqa: Derivative-free optimization algorithms by quadratic
-approximation},
-  author = {Douglas Bates and Katharine M. Mullen and John C. Nash and Ravi Varadhan},
-  year = {2014},
-  note = {R package version 1.2.4},
-  url = {http://CRAN.R-project.org/package=minqa},
-}
-\end{Soutput}
-\begin{Soutput}
-@Article{,
-  title = {The NLopt nonlinear-optimization package},
-  author = {Steven G. Johnson},
-  year = {?},
-  journal = {?},
-  volume = {?},
-  number = {?},
-  pages = {?},
-}
-\end{Soutput}
-\end{Schunk}
+Bates D, Maechler M, Bolker B and Walker S (2014). _lme4: Linear mixed-effects models using Eigen
+  and S4_. R package version 1.1-7, <URL: http://CRAN.R-project.org/package=lme4>.
+
+John C. Nash, Ravi Varadhan (2011). Unifying Optimization Algorithms to Aid Software System
+  Users: optimx for R. Journal of Statistical Software, 43(9), 1-14. URL
+  http://www.jstatsoft.org/v43/i09/.
+
+Jeroen Ooms (2014). The jsonlite Package: A Practical and Consistent Mapping Between JSON Data
+  and R Objects. arXiv:1403.2805 [stat.CO] URL http://arxiv.org/abs/1403.2805.
+
+Steven G. Johnson, The NLopt nonlinear-optimization package, http://ab-initio.mit.edu/nlopt
+
+(For the nloptr package they want us to go through and cite each algorithm we used)
 
